### 2024-11-05 README.Rmd
Initial commit 2
more functions, more documentation, passes check

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# embeddingplyr: Tools for Working With Text Embeddings

<!-- badges: start -->
<!-- badges: end -->

## About

Common operations with word and text embeddings within a 'tidyverse'/'quanteda' workflow, as demonstrated in [Data Science for Psychology: Natural Language](http://ds4psych.com). Includes simple functions for calculating common similarity metrics, as well as higher level functions for loading pretrained word embedding models (e.g. 'GloVe'), applying them to words, aggregating to produce text embeddings, and reducing dimensionality.

## Installation

You can install the development version of embeddingplyr from [GitHub](https://github.com/) with:

``` r
# devtools package required to install quanteda from Github 
remotes::install_github("rimonim/embeddingplyr") 
```

### 2024-11-05 R/data.R

@@ -1,13 +1,12 @@
-#' 100-d GloVe Model Pretrained on Twitter (subset)
+#' 25-d GloVe Model Pretrained on Twitter (subset)
 #'
-#' An embeddings object containing a subset of glove.twitter.100d pretrained by
-#' Pennington et al. (2014) on 2B Tweets. The subset includes only embeddings for
-#' common English words, as defined by [Peter Norvig's compilation of the 1/3
-#' million most frequent English words](https://norvig.com/ngrams/count_1w.txt),
+#' An embeddings object containing a subset of glove.twitter.25d pretrained by
+#' Pennington et al. (2014) on 2B Tweets. This subset includes only embeddings for
+#' the 14,000 most common English words, as defined by [Peter Norvig's compilation](https://norvig.com/ngrams/count_1w.txt),
 #' derived from the Google Web Trillion Word Corpus.
 #' Full pretrained GloVe models are available for download at https://nlp.stanford.edu/projects/glove/
 #'
 #' @references Pennington, J., Socher, R., and Manning, C. D. (2014).
 #'   GloVe: Global Vectors for Word Representation. https://nlp.stanford.edu/projects/glove/
-#' @keywords data glove_twitter_100d
-"glove_twitter_100d"
+#' @keywords data glove_twitter_25d
+"glove_twitter_25d"

### 2024-11-05 R/embeddings.R

@@ -4,67 +4,68 @@
 #' eligible input object into an embeddings object, or check whether an object
 #' is an embeddings object.
 #'
-#' @param mat a matrix
+#' @param x a numeric matrix or dataframe in which the only non-numeric column is specified by `id_col`
+#' @param ... additional arguments to be passed to class-specific methods
 #' @examples
-#'    random_mat <- matrix(
-#'       sample(1:10, 20, replace = TRUE),
-#'       nrow = 2,
-#'       dimnames = list(c("happy", "sad"))
-#'       )
-#'    as.embeddings(random_mat)
+#' random_mat <- matrix(
+#'   sample(1:10, 20, replace = TRUE),
+#'   nrow = 2,
+#'   dimnames = list(c("happy", "sad"))
+#'   )
+#' as.embeddings(random_mat)
 
+#' @rdname embeddings
 #' @export
-as.embeddings <- function(x) {
+as.embeddings <- function(x, ...) {
   UseMethod("as.embeddings")
 }
 
-#' @export
-as.embeddings.default <- function(x) {
-  if(!any(class(x) %in% c("embeddings", "matrix", "Matrix", "data.frame"))){stop(paste(class(x),collapse = "/"), " object cannot be coerced to embeddings.")}
-}
-
 #' @noRd
-#' @method as.embeddings embeddings
 #' @export
-as.embeddings.embeddings <- function(x) {
+as.embeddings.default <- function(x, ...) {
+  if(!any(class(x) %in% c("embeddings", "matrix", "Matrix", "data.frame"))){stop(paste(class(x),collapse = "/"), " object cannot be coerced to embeddings.")}
+  if (is.null(colnames(x))){
+    colnames(x) <- paste0("dim_", 1:ncol(x))
+  }
+  class(x) <- c("embeddings", "matrix", "array")
   x
 }
 
 #' @noRd
 #' @method as.embeddings matrix
 #' @export
-as.embeddings.matrix <- function(x){
+as.embeddings.matrix <- function(x, ...){
   if (!is.numeric(x)) {
     stop("Input is not numeric")
   }
-  class(x) <- c("embeddings", "matrix", "array")
-  x
+  as.embeddings.default(x)
 }
 
 #' @noRd
 #' @method as.embeddings Matrix
 #' @export
-as.embeddings.Matrix <- function(x){
+as.embeddings.Matrix <- function(x, ...){
   if (!is.numeric(x)) {
     stop("Input is not numeric")
   }
[…]

### 2024-11-05 R/find_nearest.R

@@ -51,13 +51,13 @@ find_nearest <- function(object, newdata,
     sims <- apply(object, 1, cos_sim, target)
     embeddings <- embeddings[rev(order(sims)),]
     if (!include_self) {embeddings <- embeddings[!(rownames(embeddings) %in% newdata),]}
-    head(embeddings, top_n)
+    utils::head(embeddings, top_n)
   }else{
     target <- lapply(newdata, function(tok){as.vector(embeddings[tok,])})
     names(target) <- newdata
     sims <- lapply(target, function(vec){apply(object, 1, cos_sim, vec)})
     embeddings <- lapply(sims, function(sim){embeddings[rev(order(sim)),]})
     if (!include_self) {embeddings <- lapply(embeddings, function(x){x[!(rownames(x) %in% newdata),]})}
-    lapply(embeddings, function(x){head(x, top_n)})
+    lapply(embeddings, function(x){utils::head(x, top_n)})
   }
 }

### 2024-11-05 R/load_embeddings.R

@@ -1,36 +1,75 @@
-#' Load Embeddings Model From File
+#' Load Embeddings Model From a Text or Binary File
+#'
+#' Reads GloVe text format, word2vec binary format, and most tabular formats (csv, tsv, etc.)
 #'
 #' @param path a file path
-#' @param dimensions the dimensionality of the embeddings. The defaults assumes
-#' that this information is included in the file name as per the standard for
-#' GloVe models.
 #'
 #' @details
-#' The file must have tokens in the first column and numbers in all other columns.
+#' If using a custom tabular format, the file must have tokens in the first column and numbers in all other columns.
 #'
 #' @section Value:
 #' An embeddings object (a numeric matrix with tokens as rownames)
 
 #' @export
-load_embeddings <- function(path, dimensions = as.numeric(str_extract(path, "[:digit:]+(?=d\\.txt)"))) {
-
-  # matrix with token embeddings
-  pretrained_mod <- data.table::fread(
-    path,
-    quote = "",
-    col.names = c("token", paste0("dim_", 1:dimensions))
-  ) |>
-    distinct(token, .keep_all = TRUE) |>
-    remove_rownames() |>
-    column_to_rownames("token")
-
-  if (!all(apply(pretrained_mod, 2, is.numeric))) {
-    stop("Input is not numeric")
+load_embeddings <- function(path) {
+  path_format <- substring(path, regexpr("\\.([[:alnum:]]+)$", path) + 1L)
+  # word2vec binary format
+  if (path_format == "bin") {
+    return(read_word2vec(path))
+  }else{
+    # read table with token embeddings
+    x <- data.table::fread(path, quote = "", showProgress = TRUE)
+    id <- as.vector(x[,1])
+    duplicates <- duplicated(id)
+    if (any(duplicates)) {
+      warning("Tokens are not unique. Removing duplicates.")
+      x <- x[!duplicates,]
+      id <- as.vector(x[,1])
+    }
+    x <- x[,-1]
+    if (!all(sapply(x, is.numeric))) {
+      stop("Input is not numeric")
+    }
+    x <- as.matrix(x)
+    # set row and column names
+    rownames(x) <- id
+    if (is.null(colnames(x))){
+      colnames(x) <- paste0("dim_", 1:ncol(x))
+    }
+    # update class
[…]

### 2024-11-05 R/predict.embeddings.R

@@ -8,18 +8,18 @@
 #' An embeddings object with a row for each item in `newdata`
 #'
 #' @examples
-#'    words <- c("happy", "sad")
+#' words <- c("happy", "sad")
 #'
-#'    fake_mod <- as.embeddings(
+#' fake_mod <- as.embeddings(
 #'     matrix(
-#'       sample(1:10, 20, replace = TRUE),
-#'       nrow = 2,
-#'       dimnames = list(c("happy", "sad"))
-#'       )
+#'         sample(1:10, 20, replace = TRUE),
+#'         nrow = 2,
+#'         dimnames = list(c("happy", "sad"))
 #'     )
+#' )
 #'
-#'    texts_embeddings <- predict(fake_mod, words)
-#'    texts_embeddings
+#' texts_embeddings <- predict(fake_mod, words)
+#' texts_embeddings
 
 #' @rdname predict.embeddings
 #' @export

### 2024-11-05 R/reduce_dimensionality.R

@@ -2,7 +2,8 @@
 #'
 #' Includes methods for dataframes (in the style of `dplyr`) embeddings objects, and matrices.
 #'
-#' @param data a dataframe with one embedding per row
+#' @param x a dataframe with one embedding per row
+#' @param ... additional parameters to be passed to class-specific methods
 #' @param cols tidyselect - columns that contain numeric embedding values
 #' @param reduce_to number of dimensions to keep
 #' @param scale perform scaling in addition to centering?
@@ -14,41 +15,42 @@
 #' An embeddings object (a numeric matrix with tokens as rownames)
 
 #' @export
-reduce_dimensionality <- function(x) {
+reduce_dimensionality <- function(x, ...) {
   UseMethod("reduce_dimensionality")
 }
 
-#' @method reduce_dimensionality data.frame
+#' @noRd
 #' @export
-reduce_dimensionality.data.frame <- function(data, cols, reduce_to, scale = FALSE){
-  in_dat <- dplyr::select(data, {{ cols }})
-  if (any(is.na(in_dat))) {
+reduce_dimensionality.default <- function(x, reduce_to, scale, ...) {
+  if (any(is.na(x))) {
     warning("Input data contains missing values. Rows dropped in output.")
-    data <- tidyr::drop_na(data, {{ cols }})
   }
-  pca <- stats::prcomp(~., data = in_dat, scale = scale, rank. = reduce_to)
-  out_dat <- as.data.frame(pca$x)
-  dplyr::bind_cols( select(data, -{{ cols }}), out_dat )
+  pca <- stats::prcomp(~., data = x, scale = scale, rank. = reduce_to)
+  out <- pca$x
+  rownames(out) <- rownames(x)
+  out
 }
 
+#' @rdname reduce_dimensionality
+#' @method reduce_dimensionality data.frame
+#' @export
+reduce_dimensionality.data.frame <- function(x, cols, reduce_to, scale = FALSE, ...){
+  in_dat <- dplyr::select(x, {{ cols }})
+  out_dat <- as.data.frame(reduce_dimensionality.default(x, reduce_to, scale))
+  dplyr::bind_cols( dplyr::select(x, -{{ cols }}), out_dat )
+}
+
+#' @rdname reduce_dimensionality
 #' @method reduce_dimensionality embeddings
 #' @export
-reduce_dimensionality.embeddings <- function(data, reduce_to, scale = FALSE){
-  if (any(is.na(data))) {
[…]

### 2024-11-05 R/sim_metrics.R

@@ -4,8 +4,7 @@
 #'
 #' @param x a numeric vector
 #' @param y a numeric vector the same length as x
-#' @param pos a numeric vector the same length as x; the positive end of the anchored vector
-#' @param neg a numeric vector the same length as x; the negative end of the anchored vector
+#' @param pos,neg a pair of numeric vectors the same length as x; the positive and negative ends of the anchored vector
 #'
 #' @details
 #' `dot_prod` gives the dot product. `cos_sim` gives the cosine similarity (i.e.
@@ -16,21 +15,23 @@
 #' [Data Science for Psychology: Natural Language, Chapter 20](https://ds4psych.com/navigating-vectorspace#sec-dimension-projection).
 #'
 #' @examples
-#'    vec1 <- c(1, 5, 2)
-#'    vec2 <- c(4, 2, 2)
-#'    vec3 <- c(1, -2, -13)
+#' vec1 <- c(1, 5, 2)
+#' vec2 <- c(4, 2, 2)
+#' vec3 <- c(1, -2, -13)
 #'
-#'    dot_prod(vec1, vec2)
-#'    cos_sim(vec1, vec2)
-#'    euc_dist(vec1, vec2)
-#'    anchored_sim(vec1, vec2, vec3)
+#' dot_prod(vec1, vec2)
+#' cos_sim(vec1, vec2)
+#' euc_dist(vec1, vec2)
+#' anchored_sim(vec1, vec2, vec3)
 
+#' @rdname sim_metrics
 #' @export
 dot_prod <- function(x, y){
   dot <- x %*% y
   as.vector(dot)
 }
 
+#' @rdname sim_metrics
 #' @export
 cos_sim <- function(x, y){
   dot <- x %*% y
@@ -39,12 +40,14 @@ cos_sim <- function(x, y){
   as.vector( dot / (normx*normy) )
 }
 
+#' @rdname sim_metrics
 #' @export
 euc_dist <- function(x, y){
   diff <- x - y
   sqrt(sum(diff^2))
 }
 
+#' @rdname sim_metrics
 #' @export
 anchored_sim <- function(x, pos, neg){
   # direction vector of the line segment

### 2024-11-05 R/textstat_embedding.R

@@ -1,37 +1,70 @@
 #' Get Text Embeddings by Averaging Word Embeddings
 #'
+#' `textstat_embedding()` takes a 'quanteda' dfm. `embed_docs()` is a more
+#' versatile function for use in a purely dplyr-style workflow, which acts directly
+#' on a column of texts.
 #'
 #' @param dfm a quanteda dfm
-#' @param model an embeddings object made by `load_embeddings()` or `as.embeddings()`
+#' @param model an embeddings object made by `load_embeddings()` or `as.embeddings()`. For `embed_docs()`, `model` can alternatively be a function that takes a character vector and outputs a dataframe with a row for each element of the input
+#' @param data 	a data frame or data frame extension (e.g. a tibble)
+#' @param text_col string. a column of texts for which to compute embeddings
+#' @param id_col optional string. column of unique document ids
+#' @param ... additional parameters to pass to `quanteda::tokens()` or to the user-specified modeling function
+#' @param .keep_all description
 #' @import quanteda
 #'
 #' @section Value:
-#' A tibble with columns `doc_id`, and `V1`, `V2`, etc.
+#' A tibble with columns `doc_id`, and `V1`, `V2`, etc. or similar. If `.keep_all = TRUE`, the new columns will appear after existing ones.
 #'
 #' @examples
-#'    texts <- c("this says one thing", "and this says another")
-#'    texts_dfm <- dfm(tokens(texts))
+#' # quanteda workflow
+#' library(quanteda)
+#' texts <- c("this says one thing", "and this says another")
+#' texts_dfm <- dfm(tokens(texts))
 #'
-#'    fake_mod <- as.embeddings(
-#'     matrix(
-#'       sample(1:10, 30, replace = TRUE),
-#'       nrow = 6,
-#'       dimnames = list(c("this", "and", "says", "one", "thing", "another"))
-#'       )
-#'     )
+#' texts_embeddings <- textstat_embedding(texts_dfm, glove_twitter_25d)
+#' texts_embeddings
 #'
-#'    texts_embeddings <- textstat_embeddings(texts_dfm, fake_mod)
-#'    texts_embeddings
+#' # dplyr workflow
+#' texts_df <- data.frame(text = texts)
+#' texts_embeddings <- texts_df |> embed_docs("text", glove_twitter_25d)
+#' texts_embeddings
 
[…]

### 2024-11-05 tests/testthat.R

@@ -1,0 +1,12 @@
+# This file is part of the standard setup for testthat.
+# It is recommended that you do not modify it.
+#
+# Where should you do additional test configuration?
+# Learn more about the roles of various files in:
+# * https://r-pkgs.org/testing-design.html#sec-tests-files-overview
+# * https://testthat.r-lib.org/articles/special-files.html
+
+library(testthat)
+library(embeddingplyr)
+
+test_check("embeddingplyr")

### 2024-11-05 tests/testthat/test-embeddings.R

@@ -1,0 +1,16 @@
+test_that("Coercion to embeddings and checking for matrix", {
+  expect_true(is.embeddings(as.embeddings(matrix(1:10, nrow = 2))))
+})
+
+test_that("Coercion to embeddings and checking for dataframe", {
+    expect_true(is.embeddings(as.embeddings(data.frame(token = c("A", "B"), x = 1:2, y = 3:4))))
+})
+
+test_that("User input token_col", {
+    expect_equal(
+        rownames(
+            as.embeddings(data.frame(letter = c("A", "B"), x = 1:2, y = 3:4), id_col = "letter")
+            ),
+        c("A", "B")
+        )
+})

### 2024-11-07 README.Rmd
printing + formatting for embeddings objects
plus a few other embeddings methods

@@
 ```
+
+## Functionality
+
+embeddingplyr is designed to facilitate the use of word and text embeddings in common data manipulation and text analysis workflows, without introducing new syntax or unfamiliar data structures.
+
+embeddingplyr is model agnostic; it can be used to work with embeddings from decontextualized models like [GloVe](https://nlp.stanford.edu/projects/glove/) and [word2vec](https://code.google.com/archive/p/word2vec/), or from contextualized models like BERT or others made available through the [text](https://r-text.org) package.
+
+### Loading Pretrained Embeddings
+
+embeddingplyr won't help you train new embedding models, but it can load embeddings from a file.
+This is especially useful for pretrained word embedding models like GloVe and word2vec.
+These models can be downloaded (generally in txt or bin format, respectively) and loaded with `load_embeddings()`.
+
+```{r, warning=FALSE, message=FALSE}
+library(embeddingplyr)
+
+glove_twitter_25d <- load_embeddings("~/Documents/data/glove/glove.twitter.27B.25d.txt")
+```
+
+The outcome is an "embeddings" object.
+An embeddings object is just a numeric matrix with tokens as rownames.
+This means that it can be easily coerced to a dataframe or tibble, while also allowing special embeddings-specific methods and functions, such as `predict.embeddings()` and `find_nearest()`:
+
+```{r}
+moral_embeddings <- predict(glove_twitter_25d, c("good", "bad"))
+moral_embeddings
+
+find_nearest(glove_twitter_25d, "dog", 5L)
+```
+
+### Similarity Metrics
+
+### Tidy Workflow
+
+### Quanteda Workflow

### 2024-11-07 R/embeddings-methods.R

@@ -1,0 +1,50 @@
+#' Formatting and Printing for Embeddings Objects
+#'
+#' Since the numerical content of embeddings is generally not very informative
+#' to the naked eye, the default formatting of matrices is often cumbersome.
+#' As such, embeddings objects print in a format that allows quick visualization
+#' of dimensionality and row names.
+#'
+#' @param x the embeddings object to be formatted
+#' @param ... arguments to be passed to methods
+#' @param n integer. How many rows should be printed? Defaults to `getOption("max.print")`
+#' @param round integer. the number of decimal places to be displayed
+#'
+#' @method format embeddings
+#' @export
+format.embeddings <- function(x, ..., n = getOption("max.print"), round = 2) {
+    screen_width <- getOption("width")
+    n <- min(n, nrow(x))
+    longest_rowname <- max(max(nchar(rownames(x)[1:n])), 3L) + 2L
+    show_cols <- floor((screen_width - longest_rowname)/6) - 2L
+    show_cols <- min(show_cols, ncol(x))
+    longest_colname <- max(nchar(colnames(x)[1:show_cols]))
+    if(longest_colname > 6){
+        show_cols <- floor((screen_width - longest_rowname)/longest_colname) - 2L
+        show_cols <- min(show_cols, ncol(x))
+    }
+    x_out <- x[1:n, 1:show_cols]
+    x_out <- round(x_out, round)
+    x_out <- apply(x_out, 2, as.character)
+    x_out[x[1:n, 1:show_cols] >= 0] <- paste0(" ", x_out[x[1:n, 1:show_cols] >= 0])
+    if(show_cols < ncol(x)){
+        x_out <- cbind(x_out, rep("...", n))
+    }
+    rownames(x_out) <- rownames(x)[1:n]
+    x_out
+}
+
+#' @noRd
+#' @method print embeddings
+#' @export
+print.embeddings <- function(x, ...) {
+    x <- as.embeddings(x)
+    args <- list(...)
+    if("n" %in% names(args)){
+        n <- args$n
+    }else{
+        n <- getOption("max.print")
+    }
+    cat("\033[38;5;247m# ", ncol(x),"-dimensional embeddings with ",nrow(x)," rows \033[0m\n",sep = "")
+    print(format(x, n = n), quote = FALSE)
+}

### 2024-11-07 R/embeddings.R

@@ -13,6 +13,7 @@
 #'   dimnames = list(c("happy", "sad"))
 #'   )
 #' as.embeddings(random_mat)
+#' as.matrix(as.embeddings(random_mat))
 
 #' @rdname embeddings
 #' @export
@@ -23,12 +24,11 @@ as.embeddings <- function(x, ...) {
 #' @noRd
 #' @export
 as.embeddings.default <- function(x, ...) {
-  if(!any(class(x) %in% c("embeddings", "matrix", "Matrix", "data.frame"))){stop(paste(class(x),collapse = "/"), " object cannot be coerced to embeddings.")}
+  if(!any(class(x) %in% c("embeddings", "matrix", "Matrix", "data.frame", "array"))){stop(paste(class(x),collapse = "/"), " object cannot be coerced to embeddings.")}
   if (is.null(colnames(x))){
     colnames(x) <- paste0("dim_", 1:ncol(x))
   }
-  class(x) <- c("embeddings", "matrix", "array")
-  x
+  structure(x, class = c("embeddings", "matrix", "array"))
 }
 
 #' @noRd
@@ -52,8 +52,11 @@ as.embeddings.Matrix <- function(x, ...){
 }
 
 #' @rdname embeddings
-#' @param id_col column to take row names from. Defaults to "token"
-#' @method as.embeddings data.frame
+#' @usage
+#' \method{as.embeddings}{data.frame}(x, id_col = "token", ...)
+#' @param x A data frame to be converted into embeddings.
+#' @param id_col Column to take row names from. Defaults to `"token"`.
+#' @param ... Additional arguments passed to or from other methods.
 #' @export
 as.embeddings.data.frame <- function(x, id_col = "token", ...){
   x <- tibble::column_to_rownames(x, id_col)
@@ -69,3 +72,39 @@ as.embeddings.data.frame <- function(x, id_col = "token", ...){
 is.embeddings <- function(x, ...){
   inherits(x, "embeddings")
 }
+
+#' @rdname embeddings
+#' @method as.matrix embeddings
+#' @export
+as.matrix.embeddings <- function(x, ...){
+  structure(x, class = c("matrix", "array"))
+}
+
+#' @noRd
+#' @method '[' embeddings
+#' @export
+'[.embeddings' = function(x, ..., drop = FALSE) {
+  out <- NextMethod('[')
[…]

### 2024-11-07 R/load_embeddings.R

@@ -33,7 +33,7 @@ load_embeddings <- function(path) {
     x <- as.matrix(x)
     # set row and column names
     rownames(x) <- id
-    if (is.null(colnames(x))){
+    if (all(colnames(x) == paste0("V",2:(ncol(x)+1L)))){
       colnames(x) <- paste0("dim_", 1:ncol(x))
     }
     # update class

### 2024-11-07 R/textstat_embedding.R

@@ -16,6 +16,7 @@
 #' @section Value:
 #' A tibble with columns `doc_id`, and `V1`, `V2`, etc. or similar. If `.keep_all = TRUE`, the new columns will appear after existing ones.
 #'
+#' @aliases embed_docs
 #' @examples
 #' # quanteda workflow
 #' library(quanteda)

### 2024-11-07 R/embeddings-methods.R
embeddings stay embeddings + fixed color printing

@@ -45,6 +45,6 @@ print.embeddings <- function(x, ...) {
     }else{
         n <- getOption("max.print")
     }
-    cat("\033[38;5;247m# ", ncol(x),"-dimensional embeddings with ",nrow(x)," rows \033[0m\n",sep = "")
+    cat(pillar::style_subtle(paste0("# ", ncol(x),"-dimensional embeddings with ",nrow(x)," rows\n")))
     print(format(x, n = n), quote = FALSE)
 }

### 2024-11-07 R/find_nearest.R

@@ -51,13 +51,13 @@ find_nearest <- function(object, newdata,
     sims <- apply(object, 1, cos_sim, target)
     embeddings <- embeddings[rev(order(sims)),]
     if (!include_self) {embeddings <- embeddings[!(rownames(embeddings) %in% newdata),]}
-    utils::head(embeddings, top_n)
+    as.embeddings(utils::head(embeddings, top_n))
   }else{
     target <- lapply(newdata, function(tok){as.vector(embeddings[tok,])})
     names(target) <- newdata
     sims <- lapply(target, function(vec){apply(object, 1, cos_sim, vec)})
     embeddings <- lapply(sims, function(sim){embeddings[rev(order(sim)),]})
     if (!include_self) {embeddings <- lapply(embeddings, function(x){x[!(rownames(x) %in% newdata),]})}
-    lapply(embeddings, function(x){utils::head(x, top_n)})
+    lapply(embeddings, function(x){as.embeddings(utils::head(x, top_n))})
   }
 }

### 2024-11-07 R/predict.embeddings.R

@@ -27,5 +27,5 @@ predict.embeddings <- function(object, newdata){
   embeddings <- as.matrix(object)
   embeddings <- rbind(embeddings, matrix(ncol = ncol(embeddings), dimnames = list("NOT_IN_DICT")))
   newdata[!(newdata %in% rownames(embeddings))] <- "NOT_IN_DICT"
-  embeddings[newdata,]
+  as.embeddings(embeddings[newdata,])
 }

### 2024-11-08 README.Rmd
embeddings stay embeddings + fixed color printing
bug fixes and examples

@@
 Includes simple functions for calculating common similarity metrics, as well as higher level functions for loading pretrained word embedding models (e.g.
-'GloVe'), applying them to words, aggregating to produce text embeddings, and reducing dimensionality.
+[GloVe](https://nlp.stanford.edu/projects/glove/)), applying them to words, aggregating to produce text embeddings, and reducing dimensionality.
 
@@
 
-embeddingplyr is model agnostic; it can be used to work with embeddings from decontextualized models like [GloVe](https://nlp.stanford.edu/projects/glove/) and [word2vec](https://code.google.com/archive/p/word2vec/), or from contextualized models like BERT or others made available through the [text](https://r-text.org) package.
+embeddingplyr is model agnostic; it can be used to work with embeddings from decontextualized models like [GloVe](https://nlp.stanford.edu/projects/glove/) and [word2vec](https://code.google.com/archive/p/word2vec/), or from contextualized models like BERT or others made available through the '[text](https://r-text.org)' package.
 
@@
 
-The outcome is an "embeddings" object.
+The outcome is an embeddings object.
 An embeddings object is just a numeric matrix with tokens as rownames.
@@
 
-find_nearest(glove_twitter_25d, "dog", 5L)
+find_nearest(glove_twitter_25d, "dog", 5L, sim_func = cos_sim)
 ```

### 2024-11-08 R/embeddings-predict.R

@@ -1,0 +1,29 @@
+#' Retrieve Token Embeddings
+#'
+#' @param object an embeddings object made by `load_embeddings()` or `as.embeddings()`
+#' @param newdata a character vector of tokens
+#' @import quanteda
+#'
+#' @section Value:
+#' Either an embeddings object with a row for each item in `newdata`, or, when
+#' `newdata` is of length 1, a named numeric vector.
+#'
+#' @examples
+#' words <- c("happy", "sad")
+#'
+#' texts_embeddings <- predict(glove_twitter_25d, words)
+#' texts_embeddings
+
+#' @rdname predict.embeddings
+#' @export
+predict.embeddings <- function(object, newdata){
+    embeddings <- as.matrix(object)
+    embeddings <- rbind(embeddings, matrix(ncol = ncol(embeddings), dimnames = list("NOT_IN_DICT")))
+    newdata[!(newdata %in% rownames(embeddings))] <- "NOT_IN_DICT"
+    out <- embeddings[newdata,]
+    if(is.vector(out)){
+        out
+    }else{
+        as.embeddings(out)
+    }
+}

### 2024-11-08 R/embeddings-print.R

@@ -26,7 +26,8 @@ format.embeddings <- function(x, ..., n = getOption("max.print"), round = 2) {
     x_out <- x[1:n, 1:show_cols]
     x_out <- round(x_out, round)
     x_out <- apply(x_out, 2, as.character)
-    x_out[x[1:n, 1:show_cols] >= 0] <- paste0(" ", x_out[x[1:n, 1:show_cols] >= 0])
+    x_out[is.na(x_out)] <- "NA"
+    x_out[substr(x_out,1,1)!="-"] <- paste0(" ", x_out[substr(x_out,1,1)!="-"])
     if(show_cols < ncol(x)){
         x_out <- cbind(x_out, rep("...", n))
     }

### 2024-11-08 R/embeddings.R

@@ -6,14 +6,17 @@
 #'
 #' @param x a numeric matrix or dataframe in which the only non-numeric column is specified by `id_col`
 #' @param ... additional arguments to be passed to class-specific methods
+#' @import tibble
 #' @examples
 #' random_mat <- matrix(
 #'   sample(1:10, 20, replace = TRUE),
 #'   nrow = 2,
 #'   dimnames = list(c("happy", "sad"))
 #'   )
-#' as.embeddings(random_mat)
-#' as.matrix(as.embeddings(random_mat))
+#' random_embeddings <- as.embeddings(random_mat)
+#' as.matrix(random_embeddings)
+#'
+#' tibble::as_tibble(random_embeddings, rownames = "token")
 
 #' @rdname embeddings
 #' @export
@@ -24,7 +27,7 @@ as.embeddings <- function(x, ...) {
 #' @noRd
 #' @export
 as.embeddings.default <- function(x, ...) {
-  if(!any(class(x) %in% c("embeddings", "matrix", "Matrix", "data.frame", "array"))){stop(paste(class(x),collapse = "/"), " object cannot be coerced to embeddings.")}
+  if(!embeddings_check(x)){stop(paste(class(x),collapse = "/"), " object cannot be coerced to embeddings.")}
   if (is.null(colnames(x))){
     colnames(x) <- paste0("dim_", 1:ncol(x))
   }
@@ -83,9 +86,9 @@ as.matrix.embeddings <- function(x, ...){
 #' @noRd
 #' @method '[' embeddings
 #' @export
-'[.embeddings' = function(x, ..., drop = FALSE) {
+'[.embeddings' <- function(x, ..., drop = FALSE) {
   out <- NextMethod('[')
-  if (any(class(out) %in% c("embeddings", "matrix", "Matrix", "data.frame", "array"))) {
+  if (embeddings_check(out)) {
     as.embeddings(out)
   }else{
     out
@@ -95,7 +98,7 @@ as.matrix.embeddings <- function(x, ...){
 #' @noRd
 #' @method 'rownames<-' embeddings
 #' @export
-'rownames<-.embeddings' = function(x, value) {
+'rownames<-.embeddings' <- function(x, value) {
   attr(x, "dimnames")[[1]] <- value
   x <- as.embeddings(x)
 }
@@ -103,8 +106,25 @@ as.matrix.embeddings <- function(x, ...){
 #' @noRd
 #' @method 'colnames<-' embeddings
 #' @export
-'colnames<-.embeddings' = function(x, value) {
[…]

### 2024-11-08 R/find_nearest.R

@@ -18,18 +18,10 @@
 #' An embeddings object or list of embeddings objects.
 #'
 #' @examples
-#'    words <- c("happy", "sad")
+#' words <- c("happy", "sad")
 #'
-#'    fake_mod <- as.embeddings(
-#'     matrix(
-#'       sample(1:10, 20, replace = TRUE),
-#'       nrow = 2,
-#'       dimnames = list(c("happy", "sad"))
-#'       )
-#'     )
-#'
-#'    texts_embeddings <- predict(fake_mod, words)
-#'    texts_embeddings
+#' find_nearest(glove_twitter_25d, words)
+#' find_nearest(glove_twitter_25d, words, each = TRUE)
 
 #' @rdname find_nearest
 #' @export
@@ -48,14 +40,14 @@ find_nearest <- function(object, newdata,
     }else{
       target <- as.vector(embeddings[newdata,])
     }
-    sims <- apply(object, 1, cos_sim, target)
+    sims <- apply(object, 1, sim_func, target)
     embeddings <- embeddings[rev(order(sims)),]
     if (!include_self) {embeddings <- embeddings[!(rownames(embeddings) %in% newdata),]}
     as.embeddings(utils::head(embeddings, top_n))
   }else{
     target <- lapply(newdata, function(tok){as.vector(embeddings[tok,])})
     names(target) <- newdata
-    sims <- lapply(target, function(vec){apply(object, 1, cos_sim, vec)})
+    sims <- lapply(target, function(vec){apply(object, 1, sim_func, vec)})
     embeddings <- lapply(sims, function(sim){embeddings[rev(order(sim)),]})
     if (!include_self) {embeddings <- lapply(embeddings, function(x){x[!(rownames(x) %in% newdata),]})}
     lapply(embeddings, function(x){as.embeddings(utils::head(x, top_n))})

### 2024-11-08 R/glove_models.R

@@ -1,3 +1,0 @@
-#' Pretrained GloVe Models
-#'
-#' Two pretrained GloVe Models,

### 2024-11-08 R/predict.embeddings.R

@@ -1,31 +1,0 @@
-#' Retrieve Token Embeddings
-#'
-#' @param object an embeddings object made by `load_embeddings()` or `as.embeddings()`
-#' @param newdata a character vector of tokens
-#' @import quanteda
-#'
-#' @section Value:
-#' An embeddings object with a row for each item in `newdata`
-#'
-#' @examples
-#' words <- c("happy", "sad")
-#'
-#' fake_mod <- as.embeddings(
-#'     matrix(
-#'         sample(1:10, 20, replace = TRUE),
-#'         nrow = 2,
-#'         dimnames = list(c("happy", "sad"))
-#'     )
-#' )
-#'
-#' texts_embeddings <- predict(fake_mod, words)
-#' texts_embeddings
-
-#' @rdname predict.embeddings
-#' @export
-predict.embeddings <- function(object, newdata){
-  embeddings <- as.matrix(object)
-  embeddings <- rbind(embeddings, matrix(ncol = ncol(embeddings), dimnames = list("NOT_IN_DICT")))
-  newdata[!(newdata %in% rownames(embeddings))] <- "NOT_IN_DICT"
-  as.embeddings(embeddings[newdata,])
-}

### 2024-11-08 R/reduce_dimensionality.R

@@ -6,26 +6,37 @@
 #' @param ... additional parameters to be passed to class-specific methods
 #' @param cols tidyselect - columns that contain numeric embedding values
 #' @param reduce_to number of dimensions to keep
-#' @param scale perform scaling in addition to centering?
+#' @param center logical. Should dimensions be shifted to be centered at zero?
+#' @param scale logical. Should dimensions be scaled to have unit variance?
 #'
 #' @details
 #' The file must have tokens in the first column and numbers in all other columns.
 #'
 #' @section Value:
 #' An embeddings object (a numeric matrix with tokens as rownames)
+#'
+#' @import tibble
+#' @examples
+#' glove_2d <- reduce_dimensionality(glove_twitter_25d, 2)
+#' glove_2d
+#'
+#' library(tibble)
+#' glove_tbl <- as_tibble(glove_twitter_25d, rownames = "token")
+#' glove_tbl_2d <- glove_tbl |> reduce_dimensionality(dim_1:dim_25, 2)
+#' glove_tbl_2d
 
 #' @export
 reduce_dimensionality <- function(x, ...) {
   UseMethod("reduce_dimensionality")
 }
 
-#' @noRd
 #' @export
-reduce_dimensionality.default <- function(x, reduce_to, scale, ...) {
+reduce_dimensionality.default <- function(x, reduce_to, center = TRUE, scale = FALSE, ...) {
+  if (!embeddings_check(x)) stop(paste(class(x),collapse = "/"), " object cannot be reduced.")
   if (any(is.na(x))) {
     warning("Input data contains missing values. Rows dropped in output.")
   }
-  pca <- stats::prcomp(~., data = x, scale = scale, rank. = reduce_to)
+  pca <- stats::prcomp(~., data = x, center = center, scale. = scale, rank. = reduce_to)
   out <- pca$x
   rownames(out) <- rownames(x)
   out
@@ -34,23 +45,16 @@ reduce_dimensionality.default <- function(x, reduce_to, scale, ...) {
 #' @rdname reduce_dimensionality
 #' @method reduce_dimensionality data.frame
 #' @export
-reduce_dimensionality.data.frame <- function(x, cols, reduce_to, scale = FALSE, ...){
+reduce_dimensionality.data.frame <- function(x, cols, reduce_to, center = TRUE, scale = FALSE, ...){
   in_dat <- dplyr::select(x, {{ cols }})
[…]

### 2024-11-08 R/vector_utilities.R

@@ -1,38 +1,54 @@
 #' Additional Vector Utilities
 #'
+#' @param points a dataframe, matrix, or embeddings object in which each row
+#' represents the coordinates of a point to be projected onto the line
 #' @param line_start a vector of representing the coordinates of the line's start
 #' @param line_end a vector representing the coordinates of the line's end
-#' @param points_df a dataframe in which each row represents the coordinates of a
-#' point to be projected onto the line
 #' @details
 #' `project_points_onto_line` calculated the projections of a group of points onto
 #' a line defined by two end points. It is useful for graphing the positions of
 #' embeddings with respect to an anchored vector.
+#'
+#' @examples
+#' emotion_words <- c("happy", "sad", "scared", "angry", "anxious", "grateful")
+#' emotion_embeddings <- predict(glove_twitter_25d, emotion_words)
+#' emotion_embeddings_2d <- reduce_dimensionality(emotion_embeddings, 2)
+#'
+#' # project points
+#' happy_vec_2d <- predict(emotion_embeddings_2d, "happy")
+#' sad_vec_2d <- predict(emotion_embeddings_2d, "sad")
+#' emotional_projections <- emotion_embeddings_2d |>
+#'   project_points_onto_line(happy_vec_2d, sad_vec_2d)
+#'
 #' @export
-project_points_onto_line <- function(line_start, line_end, points_df) {
+project_points_onto_line <- function(points, line_start, line_end) {
 
-  # Calculate the direction vector of the line segment
+  # direction vector of the line segment
   line_direction <- line_end - line_start
 
-  # Calculate the vector from the starting point of the line to the points
-  vectors_to_points <- t(t(points_df) - line_start)
+  # vector from the starting point of the line to the points
+  vectors_to_points <- t(t(points) - line_start)
 
-  # Calculate the dot product of the vectors
+  # dot product
   dot_products <- rowSums(vectors_to_points %*% line_direction)
 
-  # Calculate the squared length of the line segment
+  # calculate t for the projection
   length_squared <- sum(line_direction^2)
-
-  # Calculate the parameter t for the projection
   t <- dot_products / length_squared
 
-  # Calculate the coordinates of the projected points
-  line_start_mat <- matrix(line_start, nrow = nrow(points_df), ncol = length(line_start), byrow = TRUE)
-  line_direction_mat <- matrix(line_direction, nrow = nrow(points_df), ncol = length(line_start), byrow = TRUE)
[…]

### 2024-11-08 README.Rmd
get_similarities + code formatting

@@
 
-Common operations with word and text embeddings within a 'tidyverse'/'quanteda' workflow, as demonstrated in [Data Science for Psychology: Natural Language](http://ds4psych.com).
+embeddingplyr enables common operations with word and text embeddings within a 'tidyverse'/'quanteda' workflow, as demonstrated in [Data Science for Psychology: Natural Language](http://ds4psych.com).
 Includes simple functions for calculating common similarity metrics, as well as higher level functions for loading pretrained word embedding models (e.g.
@@
 
+Functions for similarity and distance metrics are as simple as possible:
+
+```{r}
+vec1 <- c(1, 5, 2)
+vec2 <- c(4, 2, 2)
+vec3 <- c(1, -2, -13)
+
+dot_prod(vec1, vec2)            # dot product
+cos_sim(vec1, vec2)             # cosine similarity
+euc_dist(vec1, vec2)            # Euclidean distance
+anchored_sim(vec1, vec2, vec3)  # projection to an anchored vector
+```
+
 ### Tidy Workflow

### 2024-11-08 R/embeddings-predict.R

@@ -17,13 +17,13 @@
 #' @rdname predict.embeddings
 #' @export
 predict.embeddings <- function(object, newdata){
-    embeddings <- as.matrix(object)
-    embeddings <- rbind(embeddings, matrix(ncol = ncol(embeddings), dimnames = list("NOT_IN_DICT")))
-    newdata[!(newdata %in% rownames(embeddings))] <- "NOT_IN_DICT"
-    out <- embeddings[newdata,]
-    if(is.vector(out)){
-        out
-    }else{
-        as.embeddings(out)
-    }
+  embeddings <- as.matrix(object)
+  embeddings <- rbind(embeddings, matrix(ncol = ncol(embeddings), dimnames = list("NOT_IN_DICT")))
+  newdata[!(newdata %in% rownames(embeddings))] <- "NOT_IN_DICT"
+  out <- embeddings[newdata,]
+  if(is.vector(out)){
+    out
+  }else{
+    as.embeddings(out)
+  }
 }

### 2024-11-08 R/embeddings-print.R

@@ -13,39 +13,39 @@
 #' @method format embeddings
 #' @export
 format.embeddings <- function(x, ..., n = getOption("max.print"), round = 2) {
-    screen_width <- getOption("width")
-    n <- min(n, nrow(x))
-    longest_rowname <- max(max(nchar(rownames(x)[1:n])), 3L) + 2L
-    show_cols <- floor((screen_width - longest_rowname)/6) - 2L
+  screen_width <- getOption("width")
+  n <- min(n, nrow(x))
+  longest_rowname <- max(max(nchar(rownames(x)[1:n])), 3L) + 2L
+  show_cols <- floor((screen_width - longest_rowname)/6) - 2L
+  show_cols <- min(show_cols, ncol(x))
+  longest_colname <- max(nchar(colnames(x)[1:show_cols]))
+  if(longest_colname > 6){
+    show_cols <- floor((screen_width - longest_rowname)/longest_colname) - 2L
     show_cols <- min(show_cols, ncol(x))
-    longest_colname <- max(nchar(colnames(x)[1:show_cols]))
-    if(longest_colname > 6){
-        show_cols <- floor((screen_width - longest_rowname)/longest_colname) - 2L
-        show_cols <- min(show_cols, ncol(x))
-    }
-    x_out <- x[1:n, 1:show_cols]
-    x_out <- round(x_out, round)
-    x_out <- apply(x_out, 2, as.character)
-    x_out[is.na(x_out)] <- "NA"
-    x_out[substr(x_out,1,1)!="-"] <- paste0(" ", x_out[substr(x_out,1,1)!="-"])
-    if(show_cols < ncol(x)){
-        x_out <- cbind(x_out, rep("...", n))
-    }
-    rownames(x_out) <- rownames(x)[1:n]
-    x_out
+  }
+  x_out <- x[1:n, 1:show_cols]
+  x_out <- round(x_out, round)
+  x_out <- apply(x_out, 2, as.character)
+  x_out[is.na(x_out)] <- "NA"
+  x_out[substr(x_out,1,1)!="-"] <- paste0(" ", x_out[substr(x_out,1,1)!="-"])
+  if(show_cols < ncol(x)){
+    x_out <- cbind(x_out, rep("...", n))
+  }
+  rownames(x_out) <- rownames(x)[1:n]
+  x_out
 }
 
 #' @noRd
 #' @method print embeddings
 #' @export
 print.embeddings <- function(x, ...) {
-    x <- as.embeddings(x)
[…]

### 2024-11-08 R/get_similarities.R

@@ -1,0 +1,95 @@
+#' Row-wise Similarity and Distance Metrics
+#'
+#' `get_similarities(df, everything(), list(sim = vec2))` is essentially
+#' equivalent to `mutate(rowwise(df), sim = cos_sim(c_across(everything()), vec2))`.
+#' Includes methods for dataframes (in the style of `dplyr`), embeddings
+#' objects, and matrices.
+#'
+#' @param x an embeddings object, matrix, or dataframe with one embedding per row
+#' @param cols tidyselect - columns that contain numeric embedding values
+#' @param y a named list of vectors with the same dimensionality as embeddings in x.
+#' Each item will result in a column in the output, showing the similarity of each
+#' embedding in x to the vector specified in y. When `sim_func = anchored_sim`,
+#' each item of y should be a list with named vectors `pos` and `neg`.
+#' @param sim_func a function that takes two vectors and outputs a scalar
+#' similarity metric. Defaults to cosine similarity. For more options see
+#' <[`Similarity and Distance Metrics`][dot_prod]>
+#' @param ... additional parameters to be passed to methods
+#' @param .keep_all logical. Keep all columns from input?
+#'
+#' @section Value:
+#' A tibble with columns `doc_id`, and similarity metrics.
+#' If `.keep_all = TRUE`, the new columns will appear after existing ones.
+#'
+#' @import tibble
+#' @examples
+#' valence_embeddings <- predict(glove_twitter_25d, c("good", "bad"))
+#' happy_vec <- predict(glove_twitter_25d, "happy")
+#' sad_vec <- predict(glove_twitter_25d, "sad")
+#'
+#' valence_embeddings |>
+#'   get_similarities(list(happy = happy_vec))
+#' valence_embeddings |>
+#'   get_similarities(
+#'     list(happy = list(pos = happy_vec, neg = sad_vec)),
+#'     anchored_sim
+#'     )
+#'
+#' valence_df <- tibble::as_tibble(valence_embeddings, rownames = "token")
+#' valence_df |> get_similarities(
+#'   dim_1:dim_25,
+#'   list(happy = happy_vec, sad = sad_vec),
+#'   .keep_all = TRUE
+#'   )
+
+#' @export
+get_similarities <- function(x, ...) {
+  UseMethod("get_similarities")
+}
+
+#' @noRd
[…]

### 2024-11-08 R/reduce_dimensionality.R

@@ -1,6 +1,6 @@
 #' Reduce the Dimensionality of Embeddings
 #'
-#' Includes methods for dataframes (in the style of `dplyr`) embeddings objects, and matrices.
+#' Includes methods for dataframes (in the style of `dplyr`), embeddings objects, and matrices.
 #'
 #' @param x a dataframe with one embedding per row
 #' @param ... additional parameters to be passed to class-specific methods

### 2024-11-08 R/textstat_embedding.R

@@ -10,7 +10,7 @@
 #' @param text_col string. a column of texts for which to compute embeddings
 #' @param id_col optional string. column of unique document ids
 #' @param ... additional parameters to pass to `quanteda::tokens()` or to the user-specified modeling function
-#' @param .keep_all description
+#' @param .keep_all logical. Keep all columns from input?
 #' @import quanteda
 #'
 #' @section Value:

### 2024-11-10 README.Rmd
optimizations + revised documentation

@@
 embeddingplyr enables common operations with word and text embeddings within a 'tidyverse'/'quanteda' workflow, as demonstrated in [Data Science for Psychology: Natural Language](http://ds4psych.com).
-Includes simple functions for calculating common similarity metrics, as well as higher level functions for loading pretrained word embedding models (e.g.
-[GloVe](https://nlp.stanford.edu/projects/glove/)), applying them to words, aggregating to produce text embeddings, and reducing dimensionality.
+It includes simple functions for calculating common similarity metrics, as well as higher level functions for loading pretrained word embedding models (e.g.
+[GloVe](https://nlp.stanford.edu/projects/glove/)), applying them to words, aggregating to produce text embeddings, reducing dimensionality, and more.
 
@@
 
-Functions for similarity and distance metrics are as simple as possible:
+Functions for similarity and distance metrics are as simple as possible; each one takes in vectors and outputs a scalar.
 
@@
 vec2 <- c(4, 2, 2)
-vec3 <- c(1, -2, -13)
+vec3 <- c(-1, -2, -13)
 
@@
 euc_dist(vec1, vec2)            						# Euclidean distance
-anchored_sim(vec1, vec2, vec3)  # projection to an anchored vector
+anchored_sim(vec1, pos = vec2, neg = vec3)  # projection to an anchored vector
 ```
 
-### Tidy Workflow
+### Example Tidy Workflow
 
-### Quanteda Workflow
+Given a tidy dataframe of texts, `embed_docs()` will generate embeddings by averaging the embeddings of words in each text (for more information on why this works well, see [Data Science for Psychology, Chapter 18](https://ds4psych.com/decontextualized-embeddings#sec-embedding-magnitude)).
+
+```{r}
+library(dplyr)
+valence_df <- tribble(
+	~id,        ~text,
+	"positive", "happy awesome cool nice",
+	"neutral",  "ok fine sure whatever",
+	"negative", "sad bad horrible angry"
+	)
+
+valence_df <- valence_df |> 
+	embed_docs("text", glove_twitter_25d, id_col = "id", .keep_all = TRUE)
+valence_df
+```
+
+To quantify how good and how intense these texts are, we can compare them to the embeddings for "good" and "intense" using `get_similarities()`.
+Note that this step requires only a dataframe or tibble with numeric columns; the embeddings can come from any source.
+
+```{r}
+good_vec <- predict(glove_twitter_25d, "good")
+intense_vec <- predict(glove_twitter_25d, "intense")
+valence_quantified <- valence_df |> 
+	get_similarities(
+		dim_1:dim_25, 
+		list(
+			good = good_vec, 
+			intense = intense_vec
+			)
+		)
+valence_quantified
+```
+### Example Quanteda Workflow
+
+### Other Functions
+
+#### Reduce Dimensionality
+
+It is sometimes useful to reduce the dimensionality of embeddings. This is done with `reduce_dimensionality()`, which by default performs PCA without column normalization.
+
+```{r}
+valence_df_2d <- valence_df |> 
+	reduce_dimensionality(dim_1:dim_25, 2)
+valence_df_2d
+```
+#### Normalize or Center
+
+#### Magnitude

### 2024-11-10 R/find_nearest.R

@@ -41,9 +41,13 @@ find_nearest <- function(object, newdata,
       target <- as.vector(embeddings[newdata,])
     }
     sims <- apply(object, 1, sim_func, target)
-    embeddings <- embeddings[rev(order(sims)),]
-    if (!include_self) {embeddings <- embeddings[!(rownames(embeddings) %in% newdata),]}
-    as.embeddings(utils::head(embeddings, top_n))
+    if (include_self) {
+      embeddings <- embeddings[order(sims, decreasing = TRUE)[1:top_n],]
+    }else{
+      embeddings <- embeddings[order(sims, decreasing = TRUE)[1:(top_n+length(newdata))],]
+      embeddings <- embeddings[!(rownames(embeddings) %in% newdata),]
+    }
+    as.embeddings(embeddings)
   }else{
     target <- lapply(newdata, function(tok){as.vector(embeddings[tok,])})
     names(target) <- newdata

### 2024-11-10 R/get_similarities.R

@@ -13,13 +13,16 @@
 #' each item of y should be a list with named vectors `pos` and `neg`.
 #' @param sim_func a function that takes two vectors and outputs a scalar
 #' similarity metric. Defaults to cosine similarity. For more options see
-#' <[`Similarity and Distance Metrics`][dot_prod]>
+#' [Similarity and Distance Metrics][sim_metrics]
 #' @param ... additional parameters to be passed to methods
-#' @param .keep_all logical. Keep all columns from input?
+#' @param .keep_all If `TRUE`, all columns from input are retained in output.
+#' If `FALSE`, only similarity metrics will be included. If `"except.embeddings"`
+#' (the default), all columns except those used to compute the similarity will be retained.
 #'
 #' @section Value:
 #' A tibble with columns `doc_id`, and similarity metrics.
-#' If `.keep_all = TRUE`, the new columns will appear after existing ones.
+#' If `.keep_all = TRUE` or `.keep_all = "except.embeddings"`, the new columns
+#' will appear after existing ones.
 #'
 #' @import tibble
 #' @examples
@@ -84,11 +87,13 @@ get_similarities.embeddings <- function(x, y, sim_func = cos_sim, ...) {
 #' @rdname get_similarities
 #' @method get_similarities data.frame
 #' @export
-get_similarities.data.frame <- function(x, cols, y, sim_func = cos_sim, ..., .keep_all = FALSE) {
+get_similarities.data.frame <- function(x, cols, y, sim_func = cos_sim, ..., .keep_all = "except.embeddings") {
   in_dat <- dplyr::select(x, {{ cols }})
   out_dat <- get_similarities.default(in_dat, y)
-  if (.keep_all) {
+  if (.keep_all == "except.embeddings") {
   	dplyr::bind_cols( dplyr::select(x, -{{ cols }}), out_dat )
+  }else if (.keep_all) {
+  	dplyr::bind_cols( x, out_dat )
   }else{
   	out_dat
   }

### 2024-11-10 R/reduce_dimensionality.R

@@ -2,18 +2,24 @@
 #'
 #' Includes methods for dataframes (in the style of `dplyr`), embeddings objects, and matrices.
 #'
-#' @param x a dataframe with one embedding per row
+#' @param x a dataframe, embeddings object, or matrix with one embedding per row
 #' @param ... additional parameters to be passed to class-specific methods
 #' @param cols tidyselect - columns that contain numeric embedding values
-#' @param reduce_to number of dimensions to keep
+#' @param reduce_to number of dimensions to keep.
+#' The value is passed to [stats::prcomp()] as `rank.`.
 #' @param center logical. Should dimensions be shifted to be centered at zero?
 #' @param scale logical. Should dimensions be scaled to have unit variance?
 #'
 #' @details
-#' The file must have tokens in the first column and numbers in all other columns.
+#' By default, `reduce_dimensionality()`, performs principle components analysis
+#' (PCA) without column normalization, and outputs the rotated data. If `center = FALSE`
+#' and `scale = FALSE`, this is equivalent to singular value decomposition (SVD),
+#' \eqn{X = U \Sigma V^{T}}, where the output columns are equal to the first
+#' `reduce_to` columns of \eqn{U \Sigma}.
 #'
 #' @section Value:
-#' An embeddings object (a numeric matrix with tokens as rownames)
+#' An object of the same class as x, with the same number of rows but fewer columns.
+#' Reduced columns in the output will be named "PC1", "PC2", etc.
 #'
 #' @import tibble
 #' @examples

### 2024-11-10 R/sim_metrics.R

@@ -13,7 +13,7 @@
 #' anchor points, where vectors aligned with `pos` are given a score of 1 and those
 #' aligned with `neg` are given a score of 0. For more on anchored vectors, see
 #' [Data Science for Psychology: Natural Language, Chapter 20](https://ds4psych.com/navigating-vectorspace#sec-dimension-projection).
-#'
+#' @aliases sim_metrics
 #' @examples
 #' vec1 <- c(1, 5, 2)
 #' vec2 <- c(4, 2, 2)

### 2024-11-10 README.Rmd
find_nearest() support new_data in various formats
+ print/format support for single-row and single column embeddings
+ embeddings coercion for numeric vectors

@@
 
-```{r}
+```{r, warning=FALSE, message=FALSE}
 library(dplyr)

### 2024-11-10 R/embeddings-predict.R

@@ -18,8 +18,8 @@
 #' @export
 predict.embeddings <- function(object, newdata){
   embeddings <- as.matrix(object)
-  embeddings <- rbind(embeddings, matrix(ncol = ncol(embeddings), dimnames = list("NOT_IN_DICT")))
-  newdata[!(newdata %in% rownames(embeddings))] <- "NOT_IN_DICT"
+  embeddings <- rbind(embeddings, matrix(ncol = ncol(embeddings), dimnames = list("EMBEDDING_NOT_FOUND")))
+  newdata[!(newdata %in% rownames(embeddings))] <- "EMBEDDING_NOT_FOUND"
   out <- embeddings[newdata,]
   if(is.vector(out)){
     out

### 2024-11-10 R/embeddings-print.R

@@ -25,7 +25,12 @@ format.embeddings <- function(x, ..., n = getOption("max.print"), round = 2) {
   }
   x_out <- x[1:n, 1:show_cols]
   x_out <- round(x_out, round)
-  x_out <- apply(x_out, 2, as.character)
+  if (is.null(dim(x_out))) {
+    x_out <- matrix(as.character(x_out), nrow = nrow(x))
+    colnames(x_out) <- colnames(x)[1:show_cols]
+  }else{
+    x_out <- apply(x_out, 2, as.character)
+  }
   x_out[is.na(x_out)] <- "NA"
   x_out[substr(x_out,1,1)!="-"] <- paste0(" ", x_out[substr(x_out,1,1)!="-"])
   if(show_cols < ncol(x)){

### 2024-11-10 R/embeddings.R

@@ -28,6 +28,9 @@ as.embeddings <- function(x, ...) {
 #' @export
 as.embeddings.default <- function(x, ...) {
   if(!embeddings_check(x)){stop(paste(class(x),collapse = "/"), " object cannot be coerced to embeddings.")}
+  if (is.null(rownames(x))){
+    rownames(x) <- paste0("doc_", 1:nrow(x))
+  }
   if (is.null(colnames(x))){
     colnames(x) <- paste0("dim_", 1:ncol(x))
   }
@@ -51,6 +54,15 @@ as.embeddings.Matrix <- function(x, ...){
   if (!is.numeric(x)) {
     stop("Input is not numeric")
   }
+  as.embeddings.default(x)
+}
+
+#' @noRd
+#' @method as.embeddings numeric
+#' @export
+as.embeddings.numeric <- function(x, ...){
+  if (!is.null(dim(x))) stop("array object cannot be coerced to embeddings.")
+  x <- matrix(x, nrow = 1)
   as.embeddings.default(x)
 }
 

### 2024-11-10 R/find_nearest.R

@@ -1,11 +1,14 @@
 #' Find Nearest Tokens in Embedding Space
 #'
 #' @param object an embeddings object made by `load_embeddings()` or `as.embeddings()`
-#' @param newdata a character vector of tokens
+#' @param newdata a character vector of tokens indexed by `object`,
+#' an embeddings object of the same dimensionality as `object`,
+#' or a numeric vector of the same dimensionality as `object`.
 #' @param top_n integer. How many nearest neighbors should be output? If `each = TRUE`,
 #' how many nearest neighbors should be output for each token?
 #' @param sim_func a function that takes two vectors and outputs a scalar
-#' similarity metric. Defaults to cosine similarity.
+#' similarity metric. Defaults to cosine similarity. For more options see
+#' [Similarity and Distance Metrics][sim_metrics].
 #' @param each logical. If `FALSE` (the default), the embeddings of newdata are
 #' averaged and the function will output an embeddings object with the `top_n`
 #' nearest tokens to the overall average embedding. If `TRUE`, the function will
@@ -31,29 +34,53 @@ find_nearest <- function(object, newdata,
                          each = FALSE,
                          include_self = TRUE){
   if (!("embeddings" %in% class(object))) {stop("`object` is not an embeddings object")}
-  newdata[!(newdata %in% rownames(object))] <- "NOT_IN_DICT"
-  if(all(newdata == "NOT_IN_DICT")){stop("None of the items in `newdata` are tokens in the embeddings object")}
-  embeddings <- rbind(object, matrix(ncol = ncol(object), dimnames = list("NOT_IN_DICT")))
-  if (!each) {
-    if(length(newdata) > 1){
-      target <- as.vector(apply(embeddings[newdata,], 2, mean, na.rm = TRUE))
+  if (is.character(newdata)) {
+    newdata_raw <- newdata
+    embedding_not_found <- !(newdata %in% rownames(object))
+    if (any(embedding_not_found)) {
+      warning(sum(embedding_not_found), " items in `newdata` are not tokens in the embeddings object.")
+      newdata[embedding_not_found] <- "EMBEDDING_NOT_FOUND"
+    }
+    if(all(newdata == "EMBEDDING_NOT_FOUND")){stop("None of the items in `newdata` are tokens in the embeddings object")}
[…]

### 2024-11-10 R/get_similarities.R

@@ -71,6 +71,7 @@ apply_func <- function(y_item, x, sim_func) {
 #' @export
 get_similarities.default <- function(x, y, sim_func = cos_sim, ...) {
 	if (!embeddings_check(x)) stop("x must be an embeddings object, matrix, or dataframe")
+	if (!is.list(y)) stop("y must be a named list")
 	out_cols <- lapply(y, apply_func, x = x, sim_func = sim_func)
 	tibble::as_tibble(out_cols)
 }

### 2024-11-10 R/sim_metrics.R

@@ -13,6 +13,8 @@
 #' anchor points, where vectors aligned with `pos` are given a score of 1 and those
 #' aligned with `neg` are given a score of 0. For more on anchored vectors, see
 #' [Data Science for Psychology: Natural Language, Chapter 20](https://ds4psych.com/navigating-vectorspace#sec-dimension-projection).
+#' Note that, for a given set of values of `x`, `anchored_sim(x, pos, neg)` will
+#' be perfectly correlated with `dot_prod(x, pos - neg)`.
 #' @aliases sim_metrics
 #' @examples
 #' vec1 <- c(1, 5, 2)

### 2024-11-10 README.Rmd
magnitude(), normalize(), and arbitrary modeling functions for embed_docs()

@@
 
-valence_df <- valence_df |> 
+valence_embeddings_df <- valence_df |> 
 	embed_docs("text", glove_twitter_25d, id_col = "id", .keep_all = TRUE)
-valence_df
+valence_embeddings_df
 ```
 
-To quantify how good and how intense these texts are, we can compare them to the embeddings for "good" and "intense" using `get_similarities()`.
-Note that this step requires only a dataframe or tibble with numeric columns; the embeddings can come from any source.
+`embed_docs()` can also be used to generate other types of embeddings.
+For example, we can use the '[text](https://r-text.org)' package to generate embeddings using any model available from Huggingface transformers.
+
+```{r, eval=FALSE}
+sbert_embeddings <- function(texts) {
+	text::textEmbed(
+	  texts,
+	  model = "sentence-transformers/all-MiniLM-L12-v2", # model name
+	  layers = -2,  # second to last layer (default)
+	  tokens_select = "[CLS]", # use only [CLS] token
+	  dim_name = FALSE,
+	  keep_token_embeddings = FALSE
+  )$texts[[1]]
+}
+
+valence_sbert_df <- valence_df |> 
+	embed_docs("text", sbert_embeddings, id_col = "id", .keep_all = TRUE)
+```
+
+To quantify how good and how intense the texts are, we can compare them to the embeddings for "good" and "intense" using `get_similarities()`.
+Note that this step requires only a dataframe,  tibble, or embeddings object with numeric columns; the embeddings can come from any source.
 
@@
 intense_vec <- predict(glove_twitter_25d, "intense")
-valence_quantified <- valence_df |> 
+valence_quantified <- valence_embeddings_df |> 
 	get_similarities(
@@
 
+```{r}
+library(quanteda)
+
+valence_corp <- corpus(valence_df, docid_field = "id")
+valence_corp
+
+valence_dfm <- valence_corp |> 
+	tokens() |> 
+	dfm()
+
+valence_embeddings_df <- valence_dfm |> 
+	textstat_embedding(glove_twitter_25d)
+valence_embeddings_df
+```
+
 ### Other Functions
@@
 ```{r}
-valence_df_2d <- valence_df |> 
+valence_df_2d <- valence_embeddings_df |> 
 	reduce_dimensionality(dim_1:dim_25, 2)
@@
 ```
-#### Normalize or Center
+#### Normalize (Scale Embeddings to the Unit Hypersphere)
+
+```{r}
+normalize(good_vec)
+
+normalize(moral_embeddings)
+
+valence_embeddings_df |> normalize_rows(dim_1:dim_25)
+```
 
 #### Magnitude
+
+```{r}
+magnitude(good_vec)
+
+magnitude(moral_embeddings)
+```

### 2024-11-10 R/get_similarities.R

@@ -1,7 +1,7 @@
 #' Row-wise Similarity and Distance Metrics
 #'
-#' `get_similarities(df, everything(), list(sim = vec2))` is essentially
-#' equivalent to `mutate(rowwise(df), sim = cos_sim(c_across(everything()), vec2))`.
+#' `get_similarities(df, col1:col2, list(sim = vec2))` is essentially
+#' equivalent to `mutate(rowwise(df), sim = cos_sim(c_across(col1:col2), vec2))`.
 #' Includes methods for dataframes (in the style of `dplyr`), embeddings
 #' objects, and matrices.
 #'

### 2024-11-10 R/magnitude.R

@@ -1,0 +1,37 @@
+#' Magnitude of Embeddings
+#'
+#' The magnitude, norm, or length of a vector is its Euclidean distance from the origin.
+#'
+#' @param x a numeric vector or embeddings object
+#' @param ... additional parameters to be passed to methods
+#'
+#' @section Value:
+#' a numeric vector with one item per embedding in `x`
+#'
+#' @examples
+#' vec <- c(1, 4, 2)
+#' magnitude(vec)
+#'
+#' valence_embeddings <- predict(glove_twitter_25d, c("good", "bad"))
+#' magnitude(valence_embeddings)
+
+#' @export
+magnitude <- function(x, ...) {
+	UseMethod("magnitude")
+}
+
+#' @export
+magnitude.default <- function(x, ...) {
+	if (!(any(class(x) %in% c("numeric", "embeddings")))) stop("x must be a numeric vector or an embeddings object")
+	out <- apply(x, 1, function(x){sqrt(sum(x^2))})
+	names(out) <- rownames(x)
+	out
+}
+
+#' @rdname magnitude
+#' @method magnitude numeric
+#' @export
+magnitude.numeric <- function(x, ...) {
+	if (!is.null(dim(x))) stop("x must be a numeric vector or an embeddings object")
+	sqrt(sum(x^2))
+}

### 2024-11-10 R/normalize.R

@@ -1,0 +1,72 @@
+#' Scale Embeddings to the Unit Hypersphere
+#'
+#' Normalize embeddings such that their magnitude is 1, while their angle from the origin is unchanged.
+#'
+#' @param x for `normalize()`, a numeric vector or embeddings object.
+#' For `normalize_rows()`, a matrix or dataframe with one embedding per row.
+#' @param cols tidyselect - columns that contain numeric embedding values
+#' @param ... additional parameters to be passed to methods
+#'
+#' @section Value:
+#' An object of the same class and dimensionality as x
+#'
+#' @import tibble
+#' @examples
+#' vec <- c(1, 4, 2)
+#' normalize(vec)
+#'
+#' valence_embeddings <- predict(glove_twitter_25d, c("good", "bad"))
+#' normalize(valence_embeddings)
+#'
+#' valence_df <- tibble::as_tibble(valence_embeddings, rownames = "token")
+#' valence_df |> normalize_rows(dim_1:dim_25)
+
+#' @export
+normalize <- function(x, ...) {
+	UseMethod("normalize")
+}
+
+#' @export
+normalize.default <- function(x, ...) {
+	if (!(any(class(x) %in% c("numeric", "embeddings")))){
+		stop("x must be a numeric vector or an embeddings object",
+				 ifelse(any(class(x) == "data.frame"), "\nFor data frames and matrices, use normalize_rows().", ""))
+	}
+	out <- apply(x, 1, function(x){x / sqrt(sum(x^2))}, simplify = FALSE)
+	out <- as.embeddings(do.call(rbind, out))
+	colnames(out) <- colnames(x)
+	out
+}
+
+#' @rdname normalize
+#' @method normalize numeric
+#' @export
+normalize.numeric <- function(x, ...) {
+	if (!is.null(dim(x))) stop("x must be a numeric vector, an embeddings object, or a dataframe with one embedding per row")
+	x / sqrt(sum(x^2))
+}
+
+#' @rdname normalize
+#' @export
+normalize_rows <- function(x, ...) {
+	UseMethod("normalize_rows")
+}
+
+#' @rdname normalize
+#' @export
+normalize_rows.default <- function(x, ...) {
+	if (!(any(class(x) %in% c("data.frame", "matrix")))) stop("x must be a dataframe or matrix")
[…]

### 2024-11-10 R/textstat_embedding.R

@@ -5,16 +5,22 @@
 #' on a column of texts.
 #'
 #' @param dfm a quanteda dfm
-#' @param model an embeddings object made by `load_embeddings()` or `as.embeddings()`. For `embed_docs()`, `model` can alternatively be a function that takes a character vector and outputs a dataframe with a row for each element of the input
+#' @param model an [embeddings][as.embeddings] object. For `embed_docs()`, `model`
+#' can alternatively be a function that takes a character vector and outputs a
+#' dataframe with a row for each element of the input
 #' @param data 	a data frame or data frame extension (e.g. a tibble)
 #' @param text_col string. a column of texts for which to compute embeddings
 #' @param id_col optional string. column of unique document ids
-#' @param ... additional parameters to pass to `quanteda::tokens()` or to the user-specified modeling function
+#' @param ... additional parameters to pass to `quanteda::tokens()` or to the
+#' user-specified modeling function
 #' @param .keep_all logical. Keep all columns from input?
+#' @param tolower logical. Convert all text to lowercase? If `model` is an
+#' embeddings object, this value is passed to [quanteda::dfm()].
 #' @import quanteda
 #'
 #' @section Value:
-#' A tibble with columns `doc_id`, and `V1`, `V2`, etc. or similar. If `.keep_all = TRUE`, the new columns will appear after existing ones.
+#' A tibble with columns `doc_id`, and `V1`, `V2`, etc. or similar.
+#' If `.keep_all = TRUE`, the new columns will appear after existing ones.
 #'
 #' @aliases embed_docs
 #' @examples
@@ -46,7 +52,8 @@ textstat_embedding <- function(dfm, model){
 #' @importFrom rlang :=
 #' @rdname textstat_embedding
 #' @export
-embed_docs <- function(data, text_col, model, id_col = NULL, ..., .keep_all = FALSE){
+embed_docs <- function(data, text_col, model, id_col = NULL, ...,
+                       .keep_all = FALSE, tolower = TRUE){
   if(is.null(id_col)){
     id_col <- "doc_id"
     data <- dplyr::mutate(data, doc_id = paste0("text", 1:nrow(data)))
@@ -57,11 +64,15 @@ embed_docs <- function(data, text_col, model, id_col = NULL, ..., .keep_all = FA
     names(new_cols)[1] <- id_col
[…]

### 2024-11-10 README.Rmd
revise README

@@
 ```{r, eval=FALSE}
+# function that takes character vector and outputs data frame
 sbert_embeddings <- function(texts) {
@@
 
-```{r}
+```{r, warning=FALSE, message=FALSE}
 library(quanteda)
 
+# corpus
 valence_corp <- corpus(valence_df, docid_field = "id")
@@
 
+# dfm
 valence_dfm <- valence_corp |> 
@@
 
+# compute embeddings
 valence_embeddings_df <- valence_dfm |> 
@@
 
+`normalize()` and `normalize_rows()` scale embeddings such that their magnitude is 1, while their angle from the origin is unchanged.
+
 ```{r}
@@
 
+The magnitude, norm, or length of a vector is its Euclidean distance from the origin.
+
 ```{r}

### 2024-11-11 README.Rmd
unit tests: embeddings

@@
 ```{r, eval=FALSE}
-# function that takes character vector and outputs data frame
+# function that takes character vector and outputs a data frame
 sbert_embeddings <- function(texts) {

### 2024-11-11 R/embeddings-print.R

@@ -15,7 +15,7 @@
 format.embeddings <- function(x, ..., n = getOption("max.print"), round = 2) {
   screen_width <- getOption("width")
   n <- min(n, nrow(x))
-  longest_rowname <- max(max(nchar(rownames(x)[1:n])), 3L) + 2L
+  longest_rowname <- max(nchar(rownames(x)[1:n]), 3L) + 2L
   show_cols <- floor((screen_width - longest_rowname)/6) - 2L
   show_cols <- min(show_cols, ncol(x))
   longest_colname <- max(nchar(colnames(x)[1:show_cols]))
@@ -52,5 +52,6 @@ print.embeddings <- function(x, ...) {
     n <- getOption("max.print")
   }
   cat(pillar::style_subtle(paste0("# ", ncol(x),"-dimensional embeddings with ",nrow(x)," rows\n")))
+  if (any(dim(x) == 0)) return()
   print(format(x, n = n), quote = FALSE)
 }

### 2024-11-11 R/embeddings.R

@@ -27,11 +27,11 @@ as.embeddings <- function(x, ...) {
 #' @noRd
 #' @export
 as.embeddings.default <- function(x, ...) {
-  if(!embeddings_check(x)){stop(paste(class(x),collapse = "/"), " object cannot be coerced to embeddings.")}
-  if (is.null(rownames(x))){
+  if(!is_matrixlike(x)){stop(paste(class(x),collapse = "/"), " object cannot be coerced to embeddings.")}
+  if (is.null(rownames(x)) && nrow(x) > 0){
     rownames(x) <- paste0("doc_", 1:nrow(x))
   }
-  if (is.null(colnames(x))){
+  if (is.null(colnames(x)) && ncol(x) > 0){
     colnames(x) <- paste0("dim_", 1:ncol(x))
   }
   structure(x, class = c("embeddings", "matrix", "array"))
@@ -51,6 +51,7 @@ as.embeddings.matrix <- function(x, ...){
 #' @method as.embeddings Matrix
 #' @export
 as.embeddings.Matrix <- function(x, ...){
+  x <- as.matrix(x)
   if (!is.numeric(x)) {
     stop("Input is not numeric")
   }
@@ -68,17 +69,19 @@ as.embeddings.numeric <- function(x, ...){
 
 #' @rdname embeddings
 #' @usage
-#' \method{as.embeddings}{data.frame}(x, id_col = "token", ...)
+#' \method{as.embeddings}{data.frame}(x, id_col = NULL, ...)
 #' @param x A data frame to be converted into embeddings.
-#' @param id_col Column to take row names from. Defaults to `"token"`.
+#' @param id_col Optional column to take row names from.
 #' @param ... Additional arguments passed to or from other methods.
 #' @export
-as.embeddings.data.frame <- function(x, id_col = "token", ...){
-  x <- tibble::column_to_rownames(x, id_col)
+as.embeddings.data.frame <- function(x, id_col = NULL, ...){
+  if (!is.null(id_col)) {
+    x <- tibble::column_to_rownames(x, id_col)
+  }
   if (!all(sapply(x, is.numeric))) {
     stop("Input contains non-numeric columns other than id_col")
   }
-  x <- as.matrix(x, dimnames = list(rownames(x), colnames(x)))
+  x <- data.matrix(x)
   as.embeddings.default(x)
 }
 
@@ -100,7 +103,7 @@ as.matrix.embeddings <- function(x, ...){
 #' @export
[…]

### 2024-11-11 R/get_similarities.R

@@ -70,7 +70,7 @@ apply_func <- function(y_item, x, sim_func) {
 
 #' @export
 get_similarities.default <- function(x, y, sim_func = cos_sim, ...) {
-	if (!embeddings_check(x)) stop("x must be an embeddings object, matrix, or dataframe")
+	if (!is_matrixlike(x)) stop("x must be an embeddings object, matrix, or dataframe")
 	if (!is.list(y)) stop("y must be a named list")
 	out_cols <- lapply(y, apply_func, x = x, sim_func = sim_func)
 	tibble::as_tibble(out_cols)

### 2024-11-11 R/reduce_dimensionality.R

@@ -38,7 +38,7 @@ reduce_dimensionality <- function(x, ...) {
 
 #' @export
 reduce_dimensionality.default <- function(x, reduce_to, center = TRUE, scale = FALSE, ...) {
-  if (!embeddings_check(x)) stop(paste(class(x),collapse = "/"), " object cannot be reduced.")
+  if (!is_matrixlike(x)) stop(paste(class(x),collapse = "/"), " object cannot be reduced.")
   if (any(is.na(x))) {
     warning("Input data contains missing values. Rows dropped in output.")
   }

### 2024-11-11 tests/testthat/test-embeddings.R

@@ -1,16 +1,186 @@
-test_that("Coercion to embeddings and checking for matrix", {
-  expect_true(is.embeddings(as.embeddings(matrix(1:10, nrow = 2))))
+test_that("as.embeddings works with numeric matrix", {
+  mat <- matrix(1:6, nrow = 2)
+  rownames(mat) <- c("token1", "token2")
+  colnames(mat) <- c("dim1", "dim2", "dim3")
+  emb <- as.embeddings(mat)
+
+  expect_s3_class(emb, "embeddings")
+  expect_equal(dim(emb), dim(mat))
+  expect_equal(rownames(emb), rownames(mat))
+  expect_equal(colnames(emb), colnames(mat))
+  expect_true(is.embeddings(emb))
 })
 
-test_that("Coercion to embeddings and checking for dataframe", {
-    expect_true(is.embeddings(as.embeddings(data.frame(token = c("A", "B"), x = 1:2, y = 3:4))))
+test_that("as.embeddings error for non-numeric matrix", {
+  mat <- matrix(c("a", "b", "c", "d"), nrow = 2)
+  expect_error(as.embeddings(mat), "Input is not numeric")
 })
 
-test_that("User input token_col", {
-    expect_equal(
-        rownames(
-            as.embeddings(data.frame(letter = c("A", "B"), x = 1:2, y = 3:4), id_col = "letter")
-            ),
-        c("A", "B")
-        )
+test_that("as.embeddings works with numeric vector", {
+  vec <- 1:5
+  emb <- as.embeddings(vec)
+
+  expect_s3_class(emb, "embeddings")
+  expect_equal(dim(emb), c(1, length(vec)))
+  expect_equal(rownames(emb), "doc_1")
+})
+
+test_that("as.embeddings.numeric error for array input", {
+  arr <- array(1:8, dim = c(2, 2, 2))
+  expect_error(as.embeddings(arr), "array object cannot be coerced to embeddings.")
+})
+
+test_that("as.embeddings works with data.frame", {
+  df <- data.frame(token = c("happy", "sad"),
+                   dim1 = c(1.0, 0.5),
+                   dim2 = c(0.5, 1.0))
+  emb <- as.embeddings(df, id_col = "token")
+
+  expect_s3_class(emb, "embeddings")
[…]

### 2024-11-17 README.Rmd
reduce_to optional for reduce_dimensionality

@@
 
-find_nearest(glove_twitter_25d, "dog", 5L, sim_func = cos_sim)
+find_nearest(glove_twitter_25d, "dog", 5L, method = "cosine")
 ```
@@
 ```
+`reduce_dimensionality()` can also be used to apply the same rotation other embeddings not used to find the principle components.
+
+```{r}
+new_embeddings <- predict(glove_twitter_25d, c("new", "strange"))
+
+# get rotation with `output_rotation = TRUE`
+valence_rotation_2d <- valence_embeddings_df |> 
+	reduce_dimensionality(dim_1:dim_25, 2, output_rotation = TRUE)
+
+# apply the same rotation to new embeddings
+new_with_valence_rotation <- new_embeddings |> 
+	reduce_dimensionality(custom_rotation = valence_rotation_2d)
+new_with_valence_rotation
+```
+
 #### Normalize (Scale Embeddings to the Unit Hypersphere)

### 2024-11-17 R/reduce_dimensionality.R

@@ -9,6 +9,10 @@
 #' The value is passed to [stats::prcomp()] as `rank.`.
 #' @param center logical. Should dimensions be shifted to be centered at zero?
 #' @param scale logical. Should dimensions be scaled to have unit variance?
+#' @param tol a value indicating a magnitude below which dimensions should be
+#' omitted. (Components are omitted if their standard deviations are less than
+#' or equal to tol times the standard deviation of the first component.) Value
+#' passed to [stats::prcomp()].
 #' @param custom_rotation optional rotation specification obtained by calling the
 #' function with `output_rotation = TRUE`. This will override `reduce_to`,
 #' `center`, and `scale`, and instead simply apply the custom rotation.
@@ -20,7 +24,7 @@
 #' (PCA) without column normalization, and outputs the rotated data. If `center = FALSE`
 #' and `scale = FALSE`, this is equivalent to singular value decomposition (SVD),
 #' \eqn{X = U \Sigma V^{T}}, where the output columns are equal to the first
-#' `reduce_to` columns of \eqn{U \Sigma}.
+#' `reduce_to` columns of \eqn{U \Sigma} that meet the criterion set by `tol`.
 #'
 #' @section Value:
 #' If `output_rotation = FALSE`, an object of the same class as x, with the same
@@ -36,6 +40,7 @@
 #'   \item `scale`: a vector with the standard deviation of each column of `x`
 #' }
 #'
+#' @importFrom stats sd
 #' @import tibble
 #' @examples
 #' glove_2d <- reduce_dimensionality(glove_twitter_25d, 2)
@@ -52,17 +57,17 @@ reduce_dimensionality <- function(x, ...) {
 }
 
 #' @export
-reduce_dimensionality.default <- function(x, reduce_to, center = TRUE, scale = FALSE, ..., custom_rotation = NULL, output_rotation = FALSE) {
+reduce_dimensionality.default <- function(x, reduce_to = NULL, center = TRUE, scale = FALSE, tol = NULL, ..., custom_rotation = NULL, output_rotation = FALSE) {
   if (!is_matrixlike(x)) stop(paste(class(x),collapse = "/"), " object cannot be reduced.")
   if (any(is.na(x))) {
     warning("Input data contains missing values. Rows dropped in output.")
   }
   if (is.null(custom_rotation)) {
[…]

### 2024-11-17 R/average_embedding.R
custom averaging methods for textstat_embedding() and embed_docs()

@@ -1,12 +1,41 @@
+#' @noRd
+make_embedding_weights <- function(x_names, w){
+	if(w[1] == "trillion_word"){
+		w <- trillion_word[x_names]
+		min_w <- min(trillion_word)
+	}else if (w[1] == "trillion_word_sif"){
+		alpha <- 0.005 # reasonable parameter for SIF
+		not_a_trillion <- 1024908267229 # total number of words processed by Google
+		alpha <- 0.005*1024908267229
+		w <- trillion_word[x_names]
+		min_w <- min(trillion_word)/(alpha + w)
+		w <- alpha/(alpha + w)
+	}else{
+		stopifnot("w must be either a numeric vector or the name of a supported weighting scheme" = is.numeric(w))
+		if (!is.null(names(w))) {
+			min_w <- min(w, na.rm = TRUE)
+			w <- w[x_names]
+		}
+	}
+	missing_w <- is.na(w)
+	if (all(missing_w)) warning("w does not include any applicable values")
+	if (any(missing_w)) warning("Replacing ", sum(missing_w), " unspecified weights with the minimum value in w", call. = FALSE)
+	w[missing_w] <- min_w
+	names(w) <- x_names
+	w
+}
+
 #' Average Embeddings
 #'
 #' Calculate the (weighted) average of multiple embeddings.
 #'
 #' @param x an embeddings object
-#' @param weights a numeric vector of weights for each row in x, `"trillion_word"`
-#' (125,000 English word frequencies from [Peter Norvig's compilation](https://norvig.com/ngrams/count_1w.txt),
-#' derived from the Google Web Trillion Word Corpus), or smooth inverse frequencies
-#' (SIF) calculated using the same list.
+#' @param w optional weighting for rows in x. This can be an unnamed
+#' numeric vector with one item per row of x, a named numeric vector of any
+#' length with names that match the row names of x, `"trillion_word"` (125,000
+#' English word frequencies from [Peter Norvig's compilation](https://norvig.com/ngrams/count_1w.txt),
+#' derived from the Google Web Trillion Word Corpus), or `"trillion_word_sif"`
+#' for smooth inverse frequencies (SIF) calculated using the same list.
 #' @param method method to use for averaging. `"mean"` (the default) is the standard
 #' arithmetic mean. `"median"` is the geometric median (also called spatial median or
[…]

### 2024-11-17 R/normalize.R

@@ -57,7 +57,7 @@ normalize_rows <- function(x, ...) {
 normalize_rows.default <- function(x, ...) {
 	if (!(any(class(x) %in% c("data.frame", "matrix")))) stop("x must be a dataframe or matrix")
 	out <- apply(x, 1, function(x){x / sqrt(sum(x^2))}, simplify = FALSE)
-	out <- as.embeddings(do.call(rbind, out))
+	out <- do.call(rbind, out)
 	colnames(out) <- colnames(x)
 	out
 }

### 2024-11-17 R/textstat_embedding.R

@@ -7,7 +7,12 @@
 #' @param dfm a quanteda dfm
 #' @param model an [embeddings][as.embeddings] object. For `embed_docs()`, `model`
 #' can alternatively be a function that takes a character vector and outputs a
-#' dataframe with a row for each element of the input
+#' dataframe with a row for each element of the input.
+#' @param w optional weighting for embeddings in `model` if `model` is an
+#' embeddings object. See [average_embedding()].
+#' @param method method to use for averaging. See [average_embedding()]. Note that
+#' `method = "median"` does not use matrix operations and may therefore be slow
+#' for datasets with many documents.
 #' @param data 	a data frame or data frame extension (e.g. a tibble)
 #' @param text_col string. a column of texts for which to compute embeddings
 #' @param id_col optional string. column of unique document ids
@@ -38,21 +43,48 @@
 #' texts_embeddings
 
 #' @export
-textstat_embedding <- function(dfm, model){
-  if(!inherits(model, "embeddings")){stop("`model` must be an embeddings object")}
+textstat_embedding <- function(dfm, model, w = NULL, method = "mean"){
+  if(!inherits(model, "embeddings")) stop("`model` must be an embeddings object")
+  if (!(method %in% c("mean", "median", "sum"))) stop("'", method, "' is not a recognized averaging method")
   feats <- quanteda::featnames(dfm)
+  # weights
+  if (!is.null(w)) {
+    w <- make_embedding_weights(rownames(model), w)
+    dfm <- suppressWarnings( quanteda::dfm_weight(dfm, weights = w) )
+  }
   # find word embeddings
   feat_embeddings <- predict.embeddings(model, feats)
-  feat_embeddings[is.na(feat_embeddings)] <- 0
-  # average word embeddings of each document
-  out_mat <- (dfm %*% feat_embeddings)/quanteda::ntoken(dfm)
-  tibble::as_tibble(as.matrix(out_mat), rownames = "doc_id")
+
+  if (method == "median") {
+    out_list <- lapply(seq_len(nrow(dfm)), function(r){
+      w <- as.numeric(dfm[r,])
+      names(w) <- feats
+      w
+      })
+    out_list <- lapply(out_list, function(w){
+      stopifnot("package 'Gmedian' is required" = requireNamespace("Gmedian", quietly = TRUE))
[…]

### 2024-11-18 README.Rmd
custom averaging methods for textstat_embedding() and embed_docs()
Easy (down)loading for GloVe, fastText, word2vec, and ConceptNet models

@@
 )
+options(embeddings.model.path = "~/Documents/data/embeddings")
 ```
@@
 
-embeddingplyr won't help you train new embedding models, but it can load embeddings from a file.
-This is especially useful for pretrained word embedding models like GloVe and word2vec.
-These models can be downloaded (generally in txt or bin format, respectively) and loaded with `load_embeddings()`.
+embeddingplyr won't help you train new embedding models, but it can load embeddings from a file or download them from online.
+This is especially useful for pretrained word embedding models like GloVe, word2vec, and fastText.
+Dozens of these models can be conveniently downloaded from online sources with `load_embeddings()`.
 
@@
 
-glove_twitter_25d <- load_embeddings("~/Documents/data/glove/glove.twitter.27B.25d.txt")
+glove_twitter_25d <- load_embeddings("glove.twitter.27B.25d")
 ```
@@
 Given a tidy dataframe of texts, `embed_docs()` will generate embeddings by averaging the embeddings of words in each text (for more information on why this works well, see [Data Science for Psychology, Chapter 18](https://ds4psych.com/decontextualized-embeddings#sec-embedding-magnitude)).
+By default, `embed_docs()` uses a simple unweighted mean, but other averaging methods are available.
 

### 2024-11-18 R/load_embeddings.R

@@ -1,8 +1,453 @@
-#' Load Embeddings Model From a Text or Binary File
+#' @noRd
+supported_models <- c(
+  glove.42B.300d = "https://huggingface.co/stanfordnlp/glove/resolve/main/glove.42B.300d.zip",
+  glove.840B.300d = "https://huggingface.co/stanfordnlp/glove/resolve/main/glove.840B.300d.zip",
+  glove.6B.50d = "https://github.com/piskvorky/gensim-data/releases/download/glove-wiki-gigaword-50/glove-wiki-gigaword-50.gz",
+  glove.6B.100d = "https://github.com/piskvorky/gensim-data/releases/download/glove-wiki-gigaword-100/glove-wiki-gigaword-100.gz",
+  glove.6B.200d = "https://github.com/piskvorky/gensim-data/releases/download/glove-wiki-gigaword-200/glove-wiki-gigaword-200.gz",
+  glove.6B.300d = "https://github.com/piskvorky/gensim-data/releases/download/glove-wiki-gigaword-300/glove-wiki-gigaword-300.gz",
+  glove.twitter.27B.25d = "https://github.com/piskvorky/gensim-data/releases/download/glove-twitter-25/glove-twitter-25.gz",
+  glove.twitter.27B.50d = "https://github.com/piskvorky/gensim-data/releases/download/glove-twitter-50/glove-twitter-50.gz",
+  glove.twitter.27B.100d = "https://github.com/piskvorky/gensim-data/releases/download/glove-twitter-100/glove-twitter-100.gz",
+  glove.twitter.27B.200d = "https://github.com/piskvorky/gensim-data/releases/download/glove-twitter-200/glove-twitter-200.gz",
+  GoogleNews.vectors.negative300 = "https://github.com/piskvorky/gensim-data/releases/download/word2vec-google-news-300/word2vec-google-news-300.gz",
+  numberbatch.19.08 = "https://conceptnet.s3.amazonaws.com/downloads/2019/numberbatch/numberbatch-19.08.txt.gz",
[…]

### 2024-11-18 R/reduce_dimensionality.R

@@ -31,8 +31,7 @@
 #' number of rows but fewer columns. Reduced columns in the output will be named
 #' "PC1", "PC2", etc.
 #'
-#' If `output_rotation = TRUE`, a list with the following
-#' components:
+#' If `output_rotation = TRUE`, a list with the following components:
 #' \itemize{
 #'   \item `rotation`: a rotation matrix
 #'   \item `center`: a vector describing how much to offset each dimension to

### 2024-11-22 R/data.R
hash tables for fast indexing

@@ -1,8 +1,9 @@
 #' 25-d GloVe Model Pretrained on Twitter (subset)
 #'
 #' An embeddings object containing a subset of glove.twitter.25d pretrained by
-#' Pennington et al. (2014) on 2B Tweets. This subset includes only embeddings for
-#' the 14,000 most common English words, as defined by [Peter Norvig's compilation](https://norvig.com/ngrams/count_1w.txt),
+#' Pennington et al. (2014) on 2B Tweets. This subset includes only embeddings
+#' included in the 12,000 most common English words, as defined by
+#' [Peter Norvig's compilation](https://norvig.com/ngrams/count_1w.txt),
 #' derived from the Google Web Trillion Word Corpus.
 #' Full pretrained GloVe models are available for download at https://nlp.stanford.edu/projects/glove/
 #'

### 2024-11-22 R/embeddings-methods.R

@@ -1,0 +1,99 @@
+#' Methods for Embeddings Objects
+#'
+#' Functions that modify the rownames or indices of embeddings objects
+#' (e.g. `unique`, `rbind`, `rownames<-` etc.) will behave as usual
+#' for matrices while automatically updating the hash table as necessary.
+#' Functions that coerce embeddings objects to other data types (e.g. `as.matrix`)
+#' will strip the hash table attribute, leaving a clean, familiar object.
+#'
+#' @param x an embeddings object
+#' @param ... for `rbind`, embeddings objects or other objects that will be coerced
+#' to embeddings. Otherwise, arguments for particular methods.
+#' @inheritParams base::rownames
+#' @inheritParams base::dimnames
+#' @inheritParams base::unique
+#' @inheritParams base::t
+#' @inheritParams base::rbind
+#' @inheritParams base::as.matrix
+#' @inheritParams tibble::as_tibble
+
+#' @rdname embeddings-methods
+#' @method `rownames<-` embeddings
+#' @usage
+#' \method{rownames}{embeddings}(x) <- value
+#' @export
+`rownames<-.embeddings` <- function(x, value) {
+	attr(x, "dimnames")[[1]] <- value
+	x <- as.embeddings(x, rowname_repair = FALSE)
+	x
+}
+
+#' @rdname embeddings-methods
+#' @method `dimnames<-` embeddings
+#' @usage
+#' \method{dimnames}{embeddings}(x) <- value
+#' @export
+`dimnames<-.embeddings` <- function(x, value) {
+	attr(x, "dimnames") <- value
+	x <- structure(x, class = c("embeddings", "matrix", "array"))
+	build_token_index(x)
+}
+
+#' @noRd
+#' @method `dim<-` embeddings
+#' @export
+`dim<-.embeddings` <- function(x, value) {
+	stop("Cannot change dimensions of an embeddings object directly.")
+}
+
+#' @rdname embeddings-methods
+#' @method unique embeddings
+#' @export
+unique.embeddings <- function(x, incomparables = FALSE, MARGIN = 1,
+															fromLast = FALSE, ...) {
+	x <- NextMethod("unique")
+	as.embeddings(x, rowname_repair = FALSE)
+}
+
+#' @rdname embeddings-methods
+#' @method t embeddings
+#' @export
+t.embeddings <- function(x) {
+	x <- NextMethod("t")
[…]

### 2024-11-22 R/embeddings-predict.R

@@ -7,6 +7,11 @@
 #' will be ignored. If `TRUE`, they will be returned as `NA`.
 #' @import quanteda
 #'
+#' @details
+#' Duplicated items in `newdata` will result in duplicated rows in the output.
+#' If an item in `newdata` matches multiple rows in `object`, the last one will
+#' be returned.
+#'
 #' @section Value:
 #' Either an embeddings object with a row for each item in `newdata`, or, when
 #' `newdata` is of length 1, a named numeric vector.
@@ -20,19 +25,16 @@
 #' @rdname predict.embeddings
 #' @export
 predict.embeddings <- function(object, newdata, .keep_missing = FALSE){
-  embedding_not_found <- !(newdata %in% rownames(object))
+  embedding_not_found <- !sapply(newdata, exists, envir = attr(object, "token_index"))
   if (any(embedding_not_found)) {
     warning(sprintf("%d items in `newdata` are not present in the embeddings object.", sum(embedding_not_found)))
   }
   if (.keep_missing) {
     out <- matrix(nrow = length(newdata), ncol = ncol(object), dimnames = list(newdata, colnames(object)))
     out[!embedding_not_found,] <- object[newdata[!embedding_not_found],]
+    out <- as.embeddings(out, rowname_repair = FALSE)
   }else{
     out <- object[newdata[!embedding_not_found],]
   }
-  if(is.vector(out)){
-    out
-  }else{
-    as.embeddings(out, .rowname_repair = FALSE)
-  }
+  out
 }

### 2024-11-22 R/embeddings-subsetting.R

@@ -1,0 +1,69 @@
+#' Extract or Subset Parts of Embeddings Objects
+#'
+#' Extraction, replacement, and subsetting nearly identically matches the behavior
+#' of matrices, with one exception: If a character item in `i` matches multiple
+#' rownames in `x`, the _last_ match will be returned.
+#'
+#' @param i row index or indices to extract or replace. Can be `numeric` or `character`.
+#' @param j column index or indeces to extract or replace. Can be `numeric` or `character`.
+#' @param drop logical. If `TRUE` (the default) and the result is one-dimensional
+#' (e.g. a single row), the output will be a (named) vector.
+#' @param value typically a numeric vector, matrix, or embeddings object.
+#' @inheritParams base::subset
+#'
+#' @details
+#' The difference between `embeddings[i,]` and `predict(embeddings, i)` is that
+#' the former will throw an error when items of `i` are not valid indices, whereas
+#' the latter will handle it gracefully (at the cost of a few more milliseconds
+#' if `i` is long).
+#'
+#' @examples
+#' glove_twitter_25d["this",]
+#' glove_twitter_25d[c("this", "that"),]
+#'
+#' glove_twitter_25d[1,]
+#' glove_twitter_25d[1:10,]
+#'
+#' glove_twitter_25d[1]
+#' glove_twitter_25d[1,1:10]
+#'
+#' duplicate_tokens <- embeddings(
+#'   1:15,
+#'   nrow = 3,
+#'   dimnames = list(c("this", "that", "this"))
+#'   )
+#' duplicate_tokens["this",]
+
+#' @rdname embeddings-subsetting
+#' @export
+#' @aliases [.embeddings
+`[.embeddings` <- function(x, i, j, drop = TRUE) {
+	if (!missing(i) && is.character(i)) {
+		i <- unlist(mget(i, envir = attr(x, "token_index")))
+	}
+	out <- NextMethod("[", i)
+	if (drop && !is_matrixlike(out)) {
+		out
+	}else{
+		as.embeddings(out, rowname_repair = FALSE)
+	}
+}
+
+#' @rdname embeddings-subsetting
+#' @export
+#' @aliases [<-.embeddings
+`[<-.embeddings` <- function(x, i, j, value) {
+	if (!missing(i) && is.character(i)) {
+		i <- unlist(mget(i, envir = attr(x, "token_index")))
+	}
+	NextMethod("[<-", i)
+}
+
[…]

### 2024-11-22 R/embeddings.R

@@ -1,14 +1,26 @@
-#' Coercion and checking functions for embeddings objects
+#' Embeddings Objects
 #'
-#' An embeddings object is a numeric matrix with rownames (generally tokens). Convert an
-#' eligible input object into an embeddings object, or check whether an object
-#' is an embeddings object.
+#' An embeddings object is a numeric matrix with fast indexing by rownames
+#' (generally tokens).
 #'
-#' @param x a numeric matrix or dataframe in which the only non-numeric column is specified by `id_col`
+#' @param x a numeric matrix or dataframe in which the only non-numeric column is
+#' specified by `id_col`, or an existing embeddings object
 #' @param ... additional arguments to be passed to or from class-specific methods
-#' @param .rowname_repair logical. If `TRUE` (the default), check that unique
+#' @param rowname_repair logical. If `TRUE` (the default), check that unique
 #' rownames are provided, and name rows "doc_1", "doc_2", etc. if not.
-#' @importFrom tibble column_to_rownames as_tibble
+#' @param rebuild_token_index logical. If `TRUE`, the hash table index will be rebuilt
+#' even when `x` is an embeddings object.
+#' @inheritParams base::matrix
+#'
+#' @details
+#' Fast row indexing is implemented using hash tables in native R environments.
+#' The "token_index" attribute of an embeddings object stores the environment
+#' that maps rownames to their corresponding indices.
+#'
+#' If `dimnames` is not supplied, `embeddings` will automatically name rows
+#' doc_1, doc_2, etc., and columns dim_1, dim_2, etc.
+#'
+#' @importFrom tibble as_tibble
 #' @examples
 #' random_mat <- matrix(
 #'   sample(1:10, 20, replace = TRUE),
@@ -16,9 +28,25 @@
 #'   dimnames = list(c("happy", "sad"))
 #'   )
 #' random_embeddings <- as.embeddings(random_mat)
-#' as.matrix(random_embeddings)
+#' is.embeddings(random_embeddings[,2:5])
 #'
 #' tibble::as_tibble(random_embeddings, rownames = "token")
+
+#' @export
[…]

### 2024-11-22 R/find_nearest.R

@@ -90,7 +90,7 @@ find_nearest <- function(object, newdata,
       target <- newdata
       available_tokens <- rownames(newdata)
     }else{
-      target <- as.embeddings(newdata, .rowname_repair = FALSE)
+      target <- as.embeddings(newdata, rowname_repair = FALSE)
       available_tokens <- rownames(target)
       target <- lapply(available_tokens, function(tok){target[tok,]})
       names(target) <- available_tokens
@@ -117,7 +117,7 @@ find_nearest <- function(object, newdata,
     if (get_similarities) {
       tibble::tibble(doc_id = rownames(object), {{method}} := sims[sims_order])
     }else{
-      as.embeddings(object, .rowname_repair = FALSE)
+      as.embeddings(object, rowname_repair = FALSE)
     }
   }else{
     sims <- lapply(target, function(vec) get_similarities(object, y = list(sim = vec), method = method, ...) )

### 2024-11-22 R/get_similarities.R

@@ -139,7 +139,6 @@ get_similarities.default <- function(x, y, method = c("cosine", "cosine_squished
 #' @method get_similarities embeddings
 #' @export
 get_similarities.embeddings <- function(x, y, method = c("cosine", "cosine_squished", "euclidean", "minkowski", "dot_prod", "anchored"), ...) {
-	x <- as.embeddings(x, .rowname_repair = FALSE)
 	out <- get_similarities.default(x, y, method, ...)
 	dplyr::bind_cols(tibble::tibble(doc_id = rownames(x)), out)
 }

### 2024-11-22 R/textstat_embedding.R

@@ -53,7 +53,7 @@ textstat_embedding <- function(dfm, model, w = NULL, method = "mean"){
     dfm <- suppressWarnings( quanteda::dfm_weight(dfm, weights = w) )
   }
   # find word embeddings
-  feat_embeddings <- predict.embeddings(model, feats)
+  feat_embeddings <- predict.embeddings(model, feats, .keep_missing = TRUE)
 
   if (method == "median") {
     out_list <- lapply(seq_len(nrow(dfm)), function(r){

### 2024-11-22 data-raw/glove_twitter_25d.R

@@ -1,0 +1,8 @@
+devtools::load_all()
+
+options(embeddings.model.path = "~/Documents/data/embeddings")
+glove_twitter_25d <- load_embeddings("glove.twitter.27B.25d")
+top_words <- trillion_word[1:12000]
+glove_twitter_25d <- predict(glove_twitter_25d, names(top_words))
+
+usethis::use_data(glove_twitter_25d, overwrite = TRUE)

### 2024-11-22 tests/testthat/test-embeddings-methods.R

@@ -1,0 +1,237 @@
+test_that("as.matrix.embeddings removes 'embeddings' class", {
+	expect_warning(emb <- as.embeddings(matrix(1:6, nrow = 2)), "unique row names not provided. Naming rows doc_1, doc_2, etc.")
+	mat2 <- as.matrix(emb)
+
+	expect_false(inherits(mat2, "embeddings"))
+	expect_true(is.matrix(mat2))
+})
+
+test_that("Setting rownames and colnames on embeddings", {
+	mat <- matrix(1:6, nrow = 2)
+	expect_warning(emb <- as.embeddings(mat), "unique row names not provided. Naming rows doc_1, doc_2, etc.")
+	rownames(emb) <- c("token1", "token2")
+	colnames(emb) <- c("dim1", "dim2", "dim3")
+
+	expect_equal(rownames(emb), c("token1", "token2"))
+	expect_equal(colnames(emb), c("dim1", "dim2", "dim3"))
+	expect_s3_class(emb, "embeddings")
+})
+
+test_that("as_tibble.embeddings converts embeddings to tibble", {
+	mat <- matrix(1:6, nrow = 2)
+	rownames(mat) <- c("token1", "token2")
+	emb <- as.embeddings(mat)
+	tbl <- tibble::as_tibble(emb, rownames = "token")
+
+	expect_s3_class(tbl, "tbl_df")
+	expect_equal(tbl$token, c("token1", "token2"))
+	expect_equal(tbl$dim_1, c(1, 2))
+})
+
+test_that("Using rbind maintains correct token_index", {
+	mat1 <- matrix(1:4, nrow = 2, dimnames = list(c("token1", "token2")))
+	embeddings1 <- as.embeddings(mat1)
+
+	mat2 <- matrix(5:8, nrow = 2, dimnames = list(c("token3", "token4")))
+	embeddings2 <- as.embeddings(mat2)
+
+	embeddings_combined <- rbind(embeddings1, embeddings2)
+
+	# Verify token_index
+	token_index <- attr(embeddings_combined, "token_index")
+	expect_equal(sort(ls(token_index)), c("token1", "token2", "token3", "token4"))
+	expect_equal(token_index[["token3"]], 3)
+})
+
+test_that("rownames<-.embeddings assigns new rownames and updates token_index", {
+	embeddings <- embeddings(1:9, nrow = 3, dimnames = list(paste0("token", 1:3), paste0("dim", 1:3)))
+
[…]

### 2024-11-22 tests/testthat/test-embeddings-subsetting.R

@@ -1,0 +1,109 @@
+test_that("Indexing embeddings retains embeddings class when multiple rows remain", {
+	mat <- matrix(1:9, nrow = 3)
+	expect_warning(emb <- as.embeddings(mat), "unique row names not provided. Naming rows doc_1, doc_2, etc.")
+	emb_sub <- emb[1:2, ]
+
+	expect_s3_class(emb_sub, "embeddings")
+	expect_equal(dim(emb_sub), c(2, 3))
+})
+
+test_that("Indexing embeddings with logical indices works correctly", {
+	emb <- embeddings(1:9, nrow = 3)
+	result <- emb[c(TRUE, FALSE, TRUE), ]
+	expect_equal(rownames(result), c("doc_1", "doc_3"))
+
+	# Length mismatch
+	expect_error(emb[c(TRUE, FALSE, TRUE, TRUE), ], "logical subscript too long")
+})
+
+test_that("Indexing embeddings returns numeric vector when appropriate", {
+	mat <- matrix(1:9, nrow = 3)
+	expect_warning(emb <- as.embeddings(mat), "unique row names not provided. Naming rows doc_1, doc_2, etc.")
+	emb_sub <- emb[1, 1]
+
+	expect_false(is.embeddings(emb_sub))
+	expect_equal(class(emb_sub), "integer")
+})
+
+test_that("Indexing embeddings to single row/column with drop = FALSE retains embeddings class", {
+	mat <- matrix(1:9, nrow = 3)
+	rownames(mat) <- c("a", "b", "c")
+	emb <- as.embeddings(mat)
+	emb_sub1 <- emb[1, , drop = FALSE]
+
+	expect_s3_class(emb_sub1, "embeddings")
+	expect_equal(dim(emb_sub1), c(1, 3))
+	expect_equal(rownames(emb_sub1), "a")
+
+	emb_sub2 <- emb[, 1, drop = FALSE]
+	expect_s3_class(emb_sub2, "embeddings")
+	expect_equal(dim(emb_sub2), c(3, 1))
+	expect_equal(colnames(emb_sub2), "dim_1")
+})
+
+test_that("Indexing embeddings on duplicate rowname returns last", {
+	emb <- embeddings(1:9, nrow = 3)
+	rownames(emb) <- c("a", "b", "a")
+	expect_equal(as.numeric(emb["a",]), c(3, 6, 9))
+})
+
+test_that("Indexing embeddings with missing i", {
+	emb <- embeddings(1:12, nrow = 3)
[…]

### 2024-11-22 tests/testthat/test-embeddings.R

@@ -1,3 +1,21 @@
+test_that("embeddings constructor works", {
+  embeddings <- embeddings(data = 1:6, nrow = 2, ncol = 3, dimnames = list(c("token1", "token2")))
+
+  expect_true(is.embeddings(embeddings))
+  expect_equal(dim(embeddings), c(2, 3))
+  expect_equal(rownames(embeddings), c("token1", "token2"))
+
+  # Verify token_index
+  token_index <- attr(embeddings, "token_index")
+  expect_equal(ls(token_index), c("token1", "token2"))
+
+  tokens <- c("token1", "token2", "token3")
+  mat <- matrix(1:9, nrow = 3, dimnames = list(tokens, paste0("dim_",1:3)))
+  emb <- embeddings(1:9, nrow = 3, dimnames = list(tokens))
+  expect_equal(emb, as.embeddings(mat))
+  expect_equal(as.matrix(emb), mat)
+})
+
 test_that("as.embeddings works with numeric matrix", {
   mat <- matrix(1:6, nrow = 2)
   rownames(mat) <- c("token1", "token2")
@@ -42,6 +60,10 @@ test_that("as.embeddings works with data.frame", {
   expect_equal(emb["happy", "dim1"], 1.0)
   expect_true(is.embeddings(emb))
 
+  # Check token_index
+  token_index <- attr(emb, "token_index")
+  expect_equal(ls(token_index), c("happy", "sad"))
+
   expect_warning(emb <- as.embeddings(df[,2:3]), "unique row names not provided. Naming rows doc_1, doc_2, etc.")
   expect_s3_class(emb, "embeddings")
   expect_equal(colnames(emb), c("dim1", "dim2"))
@@ -60,71 +82,23 @@ test_that("as.embeddings error for id_col not in data.frame", {
   expect_error(as.embeddings(df, id_col = "token"), "undefined columns selected")
 })
 
-test_that("as.embeddings keeps duplicates when .rowname_repair = TRUE", {
+test_that("as.embeddings keeps duplicates when rowname_repair = TRUE", {
   mat <- matrix(1:6, nrow = 3, dimnames = list(c("this", "that", "this")))
-  emb <- as.embeddings(mat, .rowname_repair = FALSE)
+  emb <- as.embeddings(mat, rowname_repair = FALSE)
   expect_equal(rownames(emb), rownames(mat))
 })
 
-test_that("as.embeddings.data.frame keeps duplicates when .rowname_repair = TRUE", {
[…]

### 2024-11-24 README.Rmd
hash tables for fast indexing
embed_tokens
+ refactoring embed_texts for consistency

@@
 The outcome is an embeddings object.
-An embeddings object is just a numeric matrix with tokens as rownames.
+An embeddings object is just a numeric matrix with fast hash table indexing by rownames (generally tokens).
 This means that it can be easily coerced to a dataframe or tibble, while also allowing special embeddings-specific methods and functions, such as `predict.embeddings()` and `find_nearest()`:
@@
 ```
+
+#### Recursive Behavior for Lists
+
+When supplied with a list of embeddings objects, `reduce_dimensionality()`, `normalize()`, and `magnitude()` will be called recursively and return a similar list.
+To give an example of when this may be useful, let's say we want to know the average magnitudes of the token embeddings in each of three sentences.
+This can be accomplished with the following code:
+
+```{r}
+texts <- c("will this text have a higher average magnitude?", 
+					 "or perhaps this one will?")
+texts_tokens_list <- embed_tokens(texts, glove_twitter_25d, output_embeddings = TRUE)
+texts_tokens_list
+
+texts_tokens_list |> 
+	magnitude() |> 
+	sapply(mean)
+```

### 2024-11-24 R/embed_tokens.R

@@ -1,0 +1,83 @@
+#' Get Embeddings of Tokens in a Text or Corpus
+#'
+#' Given a character vector or 'quanteda' object (tokens, dfm, or corpus)  and a
+#' word embeddings model in the form of an [embeddings] object, [embed_tokens()]
+#' returns the embedding for each token.
+#'
+#' @param x a character vector of texts, a data frame, a 'quanteda'
+#' [tokens][quanteda::tokens] object, or a 'quanteda' [corpus][quanteda::corpus]
+#' @param text_col string. a column of texts to be tokenized and converted to embeddings
+#' @param model a word embeddings model in the form of an [embeddings] object
+#' @param id_col optional string. column of unique document ids
+#' @param ... additional parameters to pass to `quanteda::tokens()`
+#' @param .keep_missing logical. What should be done about tokens in `x` that
+#' are not present in `model`? If `FALSE` (the default), they will be ignored.
+#' If `TRUE`, they will be returned as `NA`.
+#' @param .keep_all logical. Keep all columns from input? Ignored if `output_embeddings = TRUE`.
+#' @param tolower logical. Convert all text to lowercase?
+#' @param output_embeddings `FALSE` (the default) returns a tibble.
+#' `TRUE` returns a list of embeddings objects. See 'Value' for details.
+#'
+#' @section Value:
+#' If `output_embeddings = FALSE`, a tibble with columns `doc_id`, `token`, and embedding
+#' dimension names. If `.keep_all = TRUE`, the new columns will appear after
+#' existing ones.
+#' If `output_embeddings = TRUE`, a named list of embeddings objects with tokens as
+#' rownames.
+#'
+#' @seealso [embed_docs()]
+#'
+#' @examples
+#' texts <- c("this says one thing", "and this says another")
+#' texts_token_embeddings <- embed_tokens(texts, glove_twitter_25d)
+#' texts_token_embeddings
+#'
+#' # quanteda workflow
+#' library(quanteda)
+#' texts_tokens <- tokens(texts)
+#' texts_token_embeddings <- embed_tokens(texts_tokens, glove_twitter_25d)
+#' texts_token_embeddings
+#'
+#' # dplyr workflow
+#' texts_df <- data.frame(text = texts)
[…]

### 2024-11-24 R/embeddings-methods.R

@@ -86,14 +86,16 @@ as.matrix.embeddings <- function(x, ...){
 	as.matrix(x)
 }
 
-
+#' @importFrom rlang :=
 #' @rdname embeddings-methods
 #' @method as_tibble embeddings
 #' @export
 as_tibble.embeddings <- function(x, ..., rownames = NULL) {
+	rn <- rownames(x) %||% NA
 	x <- as.data.frame(x)
 	if (!is.null(rownames)) {
-		x <- tibble::rownames_to_column(x, var = rownames)
+		x <- dplyr::mutate(x, !!rownames := rn)
+		x <- dplyr::relocate(x, tidyselect::all_of(rownames))
 	}
 	tibble::as_tibble(x, ...)
 }

### 2024-11-24 R/textstat_embedding.R

@@ -1,8 +1,8 @@
 #' Get Text Embeddings by Averaging Word Embeddings
 #'
 #' `textstat_embedding()` takes a 'quanteda' dfm. `embed_docs()` is a more
-#' versatile function for use in a purely dplyr-style workflow, which acts directly
-#' on a column of texts.
+#' versatile function for which acts directly on either a character vector or
+#' a column of texts in a dataframe.
 #'
 #' @param dfm a quanteda dfm
 #' @param model an [embeddings] object. For `embed_docs()`, `model`
@@ -13,29 +13,34 @@
 #' @param method method to use for averaging. See [average_embedding()]. Note that
 #' `method = "median"` does not use matrix operations and may therefore be slow
 #' for datasets with many documents.
-#' @param data 	a data frame or data frame extension (e.g. a tibble)
+#' @param x a character vector, a data frame, or data frame extension (e.g. a tibble)
 #' @param text_col string. a column of texts for which to compute embeddings
 #' @param id_col optional string. column of unique document ids
 #' @param ... additional parameters to pass to `quanteda::tokens()` or to the
 #' user-specified modeling function
-#' @param .keep_all logical. Keep all columns from input? Ignored if `output = "tibble"`.
+#' @param .keep_all logical. Keep all columns from input? Ignored if `output_embeddings = TRUE`.
 #' @param tolower logical. Convert all text to lowercase? If `model` is an
 #' embeddings object, this value is passed to [quanteda::dfm()].
-#' @param output `"tibble"` (the default) returns a tibble. `"embeddings"`
-#' returns an embeddings object. See 'Value' for details.
+#' @param output_embeddings `FALSE` (the default) returns a tibble.
+#' `TRUE` returns an embeddings object. See 'Value' for details.
 #' @import quanteda
 #'
 #' @section Value:
-#' If `output = "tibble"`, a tibble with columns `doc_id`, and `dim_1`, `dim_2`,
+#' If `output_embeddings = FALSE`, a tibble with columns `doc_id`, and `dim_1`, `dim_2`,
 #' etc. or similar. If `.keep_all = TRUE`, the new columns will appear after
 #' existing ones.
-#' If `output = "embeddings"`, an embeddings object with document ids as rownames.
[…]

### 2024-11-26 README.Rmd
Update README

--- 2024-11-18 README.Rmd
@@
 ```
+
+Whereas indexing a regular matrix by rownames gets slower as the number of rows increases, embedingplyr's hash table indexing means that token embeddings can be retrieved in milliseconds even from models with millions of rows.
 

### 2024-11-28 README.Rmd
reorganize sim_metrics for matrices

@@
 
-To quantify how good and how intense the texts are, we can compare them to the embeddings for "good" and "intense" using `get_similarities()`.
+To quantify how good and how intense the texts are, we can compare them to the embeddings for "good" and "intense" using `get_sims()`.
 Note that this step requires only a dataframe,  tibble, or embeddings object with numeric columns; the embeddings can come from any source.
@@
 valence_quantified <- valence_embeddings_df |> 
-	get_similarities(
+	get_sims(
 		dim_1:dim_25, 

### 2024-11-28 R/find_nearest.R

@@ -8,8 +8,8 @@
 #' how many nearest neighbors should be output for each token?
 #' @param method either the name of a method to compute similarity or distance,
 #' or a function that takes two vectors and outputs a scalar, like those listed
-#' in [Similarity and Distance Metrics][sim_metrics]. The value is passed to [get_similarities()].
-#' @param ... additional parameters to be passed to method function by way of [get_similarities()]
+#' in [Similarity and Distance Metrics][sim_metrics]. The value is passed to [get_sims()].
+#' @param ... additional parameters to be passed to method function by way of [get_sims()]
 #' @param each logical. If `FALSE` (the default), the embeddings of newdata are
 #' averaged and the function will output an embeddings object with the `top_n`
 #' nearest tokens to the overall average embedding. If `TRUE`, the function will
@@ -21,13 +21,13 @@
 #' the embeddings with the highest similarity scores are output. If `FALSE`
 #' (the default for `"euclidean"` and `"minkowski"` methods), the embeddings with
 #' the lowest distance scores are output.
-#' @param get_similarities logical. If `FALSE` (the default), output an embeddings
+#' @param get_sims logical. If `FALSE` (the default), output an embeddings
 #' object or list of embeddings objects. If `TRUE`, output a tibble with similarity
 #' or distance scores for each document.
 #'
 #' @section Value:
-#' If `get_similarities = FALSE` (the default), an embeddings object or list of
-#' embeddings objects.  If `get_similarities = TRUE`, a tibble or list of tibbles
+#' If `get_sims = FALSE` (the default), an embeddings object or list of
+#' embeddings objects.  If `get_sims = TRUE`, a tibble or list of tibbles
 #' with similarity or distance scores for each document, with columns `doc_id`
 #' and the name of the requested `method`.
 #'
@@ -52,7 +52,7 @@ find_nearest <- function(object, newdata,
                          each = FALSE,
                          include_self = TRUE,
                          decreasing = NULL,
-                         get_similarities = FALSE){
+                         get_sims = FALSE){
   if (!inherits(object, "embeddings")) {stop("`object` is not an embeddings object")}
   if (any(is.na(object))) stop("`object` contains NA values")
   # similarity or distance?
@@ -104,7 +104,7 @@ find_nearest <- function(object, newdata,
     if (length(available_tokens) > 1) {
[…]

### 2024-11-28 R/get_sims.R

@@ -1,6 +1,6 @@
 #' Row-wise Similarity and Distance Metrics
 #'
-#' `get_similarities(df, col1:col2, list(sim = vec2))` is essentially
+#' `get_sims(df, col1:col2, list(sim = vec2))` is essentially
 #' equivalent to `mutate(rowwise(df), sim = cos_sim(c_across(col1:col2), vec2))`.
 #' Includes methods for dataframes (in the style of `dplyr`), embeddings
 #' objects, and matrices.
@@ -50,20 +50,20 @@
 #' sad_vec <- predict(glove_twitter_25d, "sad")
 #'
 #' valence_embeddings |>
-#'   get_similarities(list(happy = happy_vec))
+#'   get_sims(list(happy = happy_vec))
 #' valence_embeddings |>
-#'   get_similarities(
+#'   get_sims(
 #'     list(happy = list(pos = happy_vec, neg = sad_vec)),
 #'     anchored_sim
 #'     )
 #' valence_embeddings |>
-#'   get_similarities(
+#'   get_sims(
 #'     list(happy = happy_vec),
 #'     method = function(x, y) sum(abs(x - y))
 #'     )
 #'
 #' valence_df <- tibble::as_tibble(valence_embeddings, rownames = "token")
-#' valence_df |> get_similarities(
+#' valence_df |> get_sims(
 #'   dim_1:dim_25,
 #'   list(happy = happy_vec, sad = sad_vec),
 #'   .keep_all = TRUE
@@ -71,16 +71,16 @@
 
 
 #' @export
-get_similarities <- function(x, ...) {
-  UseMethod("get_similarities")
+get_sims <- function(x, ...) {
+  UseMethod("get_sims")
 }
 
 #' @noRd
 #' @include sim_metrics.R
-sim_matrix_list <- list(
-	"cosine" = cos_sim_matrix, "cosine_squished" = cos_sim_squished_matrix,
-	"euclidean" = euc_dist_matrix, "minkowski" = minkowski_dist_matrix,
-	"dot_prod" = dot_prod_matrix, "anchored" = anchored_sim_matrix
+sim_mat_vec_list <- list(
+	"cosine" = cos_sim_mat_vec, "cosine_squished" = cos_sim_squished_mat_vec,
+	"euclidean" = euc_dist_mat_vec, "minkowski" = minkowski_dist_mat_vec,
+	"dot_prod" = dot_prod_mat_vec, "anchored" = anchored_sim_mat_vec
 )
 
[…]

### 2024-11-28 R/sim_metrics.R

@@ -83,21 +83,25 @@ anchored_sim <- function(x, pos, neg){
   as.numeric(proj) / anchored_norm
 }
 
-#' Faster Similarity and Distance Metrics With Matrix Operations
-#' @rdname sim_metrics_matrix
+#' Similarity and Distance Between A Matrix and a Vector
+#'
+#' These functions are equivalent to calling [vector similarity metrics][sim_metrics]
+#' on each row of a matrix, but use matrix operations and are therefore much faster.
+#'
+#' @rdname sim_metrics_mat_vec
 #' @param x a numeric matrix or embeddings object
 #' @param y a numeric vector of length `ncol(x)`
 #' @section Value:
 #' A named numeric vector of length `nrow(x)`
 #' @keywords internal
-dot_prod_matrix <- function(x, y) {
+dot_prod_mat_vec <- function(x, y) {
   stopifnot("x and y must have the same number of dimensions" = ncol(x) == length(y))
   x %*% y
 }
 
-#' @rdname sim_metrics_matrix
+#' @rdname sim_metrics_mat_vec
 #' @keywords internal
-cos_sim_matrix <- function(x, y) {
+cos_sim_mat_vec <- function(x, y) {
   stopifnot("x and y must have the same number of dimensions" = ncol(x) == length(y))
   normx <- sqrt(rowSums(x^2))
   normy <- sqrt(sum(y^2))
@@ -107,26 +111,25 @@ cos_sim_matrix <- function(x, y) {
   as.numeric(sims)
 }
 
-#' @rdname sim_metrics_matrix
+#' @rdname sim_metrics_mat_vec
 #' @keywords internal
-cos_sim_squished_matrix <- function(x, y) {
-  stopifnot("x and y must have the same number of dimensions" = ncol(x) == length(y))
-  cos <- cos_sim_matrix(x, y)
+cos_sim_squished_mat_vec <- function(x, y) {
+  cos <- cos_sim_mat_vec(x, y)
   cos*0.5 + 0.5
 }
 
-#' @rdname sim_metrics_matrix
+#' @rdname sim_metrics_mat_vec
 #' @keywords internal
-euc_dist_matrix <- function(x, y) {
+euc_dist_mat_vec <- function(x, y) {
   stopifnot("x and y must have the same number of dimensions" = ncol(x) == length(y))
[…]

### 2024-11-28 tests/testthat/test-find_nearest.R

@@ -72,14 +72,14 @@ test_that("find_nearest works with include_self = FALSE", {
 	expect_false("token1" %in% rownames(result))
 })
 
-test_that("find_nearest returns similarities when get_similarities = TRUE", {
+test_that("find_nearest returns similarities when get_sims = TRUE", {
 	embeddings_matrix <- matrix(runif(100), nrow = 10, dimnames = list(paste0("token", 1:10)))
 	embeddings <- as.embeddings(embeddings_matrix)
 
 	rownames(embeddings) <- paste0("token", 1:10)
 	newdata <- "token1"
 
-	result <- find_nearest(embeddings, newdata, top_n = 3, get_similarities = TRUE)
+	result <- find_nearest(embeddings, newdata, top_n = 3, get_sims = TRUE)
 
 	expect_true(is_tibble(result))
 	expect_equal(nrow(result), 3)
@@ -194,13 +194,13 @@ test_that("find_nearest errors when object is not an embeddings object", {
 	)
 })
 
-test_that("find_nearest works with get_similarities = TRUE and each = TRUE", {
+test_that("find_nearest works with get_sims = TRUE and each = TRUE", {
 	embeddings_matrix <- matrix(runif(100), nrow = 10, dimnames = list(paste0("token", 1:10)))
 	embeddings <- as.embeddings(embeddings_matrix)
 
 	newdata <- c("token1", "token2")
 
-	result <- find_nearest(embeddings, newdata, top_n = 3, each = TRUE, get_similarities = TRUE)
+	result <- find_nearest(embeddings, newdata, top_n = 3, each = TRUE, get_sims = TRUE)
 
 	expect_true(is.list(result))
 	expect_equal(length(result), length(newdata))

### 2024-11-28 tests/testthat/test-get_sims.R

@@ -1,4 +1,4 @@
-test_that("get_similarities.default computes similarities correctly with method 'cosine'", {
+test_that("get_sims.default computes similarities correctly with method 'cosine'", {
 	# Create a sample matrix x
 	x <- matrix(runif(30), nrow = 10, ncol = 3)
 	# Create a named list y with one vector
@@ -6,7 +6,7 @@ test_that("get_similarities.default computes similarities correctly with method 
 	y <- list(similarity = y_vec)
 
 	# Call the function
-	result <- get_similarities.default(x, y, method = "cosine")
+	result <- get_sims.default(x, y, method = "cosine")
 
 	# Check that result is a tibble
 	expect_true(is_tibble(result))
@@ -14,27 +14,27 @@ test_that("get_similarities.default computes similarities correctly with method 
 	expect_equal(nrow(result), nrow(x))
 	expect_equal(ncol(result), length(y))
 
-	# Compute expected values using cos_sim_matrix
-	expected_values <- cos_sim_matrix(x, y_vec)
+	# Compute expected values using cos_sim_mat_vec
+	expected_values <- cos_sim_mat_vec(x, y_vec)
 	# Compare the computed similarities
 	expect_equal(result$similarity, expected_values)
 })
 
-test_that("get_similarities.default handles multiple y vectors", {
+test_that("get_sims.default handles multiple y vectors", {
 	x <- matrix(runif(30), nrow = 10, ncol = 3)
 	y_vec1 <- runif(3)
 	y_vec2 <- runif(3)
 	y <- list(sim1 = y_vec1, sim2 = y_vec2)
 
-	result <- get_similarities.default(x, y, method = "cosine")
+	result <- get_sims.default(x, y, method = "cosine")
 
 	expect_true(is_tibble(result))
 	expect_equal(ncol(result), length(y))
-	expect_equal(result$sim1, cos_sim_matrix(x, y_vec1))
-	expect_equal(result$sim2, cos_sim_matrix(x, y_vec2))
+	expect_equal(result$sim1, cos_sim_mat_vec(x, y_vec1))
+	expect_equal(result$sim2, cos_sim_mat_vec(x, y_vec2))
 })
 
-test_that("get_similarities.default handles custom method functions", {
+test_that("get_sims.default handles custom method functions", {
 	x <- matrix(runif(30), nrow = 10, ncol = 3)
 	y_vec <- runif(3)
 	y <- list(custom_sim = y_vec)
[…]

### 2024-11-28 tests/testthat/test-sim_metrics.R

@@ -148,131 +148,131 @@ test_that("anchored_sim returns error for mismatched dimensions", {
 	expect_error(anchored_sim(x, pos, neg), "x, pos, and neg must have the same number of dimensions")
 })
 
-test_that("dot_prod_matrix computes dot product for each row", {
+test_that("dot_prod_mat_vec computes dot product for each row", {
 	x <- matrix(1:6, nrow = 2)
 	y <- c(1, 1, 1)
 	expected <- x %*% y
-	result <- dot_prod_matrix(x, y)
+	result <- dot_prod_mat_vec(x, y)
 	expect_equal(result, expected)
 })
 
-test_that("dot_prod_matrix handles negative values", {
+test_that("dot_prod_mat_vec handles negative values", {
 	x <- matrix(c(-1, -2, -3, -4, -5, -6), nrow = 2)
 	y <- c(1, 2, 3)
 	expected <- x %*% y
-	result <- dot_prod_matrix(x, y)
+	result <- dot_prod_mat_vec(x, y)
 	expect_equal(result, expected)
 })
 
-test_that("dot_prod_matrix returns error for mismatched dimensions", {
+test_that("dot_prod_mat_vec returns error for mismatched dimensions", {
 	x <- matrix(1:6, nrow = 2)
 	y <- c(1, 1)
-	expect_error(dot_prod_matrix(x, y), "x and y must have the same number of dimensions")
+	expect_error(dot_prod_mat_vec(x, y), "x and y must have the same number of dimensions")
 })
 
-test_that("cos_sim_matrix computes cosine similarity for each row", {
+test_that("cos_sim_mat_vec computes cosine similarity for each row", {
 	x <- matrix(c(1, 0, 0, 1, 1, 0), nrow = 2, byrow = TRUE)
 	y <- c(1, 0, 0)
 	expected <- c(1, sqrt(1/2))
-	result <- cos_sim_matrix(x, y)
+	result <- cos_sim_mat_vec(x, y)
 	expect_equal(result, expected)
 })
 
-test_that("cos_sim_matrix handles zero vectors", {
+test_that("cos_sim_mat_vec handles zero vectors", {
 	x <- matrix(c(0, 0, 0, 1, 1, 1), nrow = 2, byrow = TRUE)
 	y <- c(1, 1, 1)
-	result <- cos_sim_matrix(x, y)
[…]

### 2024-12-02 README.Rmd
change package name

@@
 
-# embeddingplyr: Tools for Working With Text Embeddings
+# embedplyr: Tools for Working With Text Embeddings
 
@@
 
-embeddingplyr enables common operations with word and text embeddings within a 'tidyverse'/'quanteda' workflow, as demonstrated in [Data Science for Psychology: Natural Language](http://ds4psych.com).
+embedplyr enables common operations with word and text embeddings within a 'tidyverse'/'quanteda' workflow, as demonstrated in [Data Science for Psychology: Natural Language](http://ds4psych.com).
 It includes simple functions for calculating common similarity metrics, as well as higher level functions for loading pretrained word embedding models (e.g.
@@
 
-You can install the development version of embeddingplyr from [GitHub](https://github.com/) with:
+You can install the development version of embedplyr from [GitHub](https://github.com/) with:
 
@@
 # devtools package required to install quanteda from Github 
-remotes::install_github("rimonim/embeddingplyr") 
+remotes::install_github("rimonim/embedplyr") 
 ```
@@
 
-embeddingplyr is designed to facilitate the use of word and text embeddings in common data manipulation and text analysis workflows, without introducing new syntax or unfamiliar data structures.
+embedplyr is designed to facilitate the use of word and text embeddings in common data manipulation and text analysis workflows, without introducing new syntax or unfamiliar data structures.
 
-embeddingplyr is model agnostic; it can be used to work with embeddings from decontextualized models like [GloVe](https://nlp.stanford.edu/projects/glove/) and [word2vec](https://code.google.com/archive/p/word2vec/), or from contextualized models like BERT or others made available through the '[text](https://r-text.org)' package.
+embedplyr is model agnostic; it can be used to work with embeddings from decontextualized models like [GloVe](https://nlp.stanford.edu/projects/glove/) and [word2vec](https://code.google.com/archive/p/word2vec/), or from contextualized models like BERT or others made available through the '[text](https://r-text.org)' package.
 
@@
 
-embeddingplyr won't help you train new embedding models, but it can load embeddings from a file or download them from online.
+embedplyr won't help you train new embedding models, but it can load embeddings from a file or download them from online.
 This is especially useful for pretrained word embedding models like GloVe, word2vec, and fastText.
@@
 ```{r, warning=FALSE, message=FALSE}
-library(embeddingplyr)
+library(embedplyr)
 

### 2024-12-02 R/load_embeddings.R

@@ -520,7 +520,7 @@ read_word2vec <- function(path, words = NULL){
   # handle for HTTP requests
   h <- curl::new_handle()
   curl::handle_setopt(h, timeout = 10000);
-  curl::handle_setheaders(h, "User-Agent" = "embeddingplyr (https://github.com/rimonim/embeddingplyr)")
+  curl::handle_setheaders(h, "User-Agent" = "embedplyr (https://github.com/rimonim/embedplyr)")
   # open binary file in read mode
   is_gz <- is_gzipped(path)
   if (is_url(path)) {
@@ -674,7 +674,7 @@ fread_filtered <- function(file, words, use_sys = TRUE, ..., timeout = 1000) {
     # handle for HTTP requests
     h <- curl::new_handle()
     curl::handle_setopt(h, timeout = 10000);
-    curl::handle_setheaders(h, "User-Agent" = "embeddingplyr (https://github.com/rimonim/embeddingplyr)")
+    curl::handle_setheaders(h, "User-Agent" = "embedplyr (https://github.com/rimonim/embedplyr)")
     # open connection
     if (is_url(file)) {
       conn <- tryCatch({

### 2024-12-02 tests/testthat.R

@@ -7,6 +7,6 @@
 # * https://testthat.r-lib.org/articles/special-files.html
 
 library(testthat)
-library(embeddingplyr)
+library(embedplyr)
 
-test_check("embeddingplyr")
+test_check("embedplyr")

### 2024-12-02 README.Rmd
logo
fix README title
update README

@@
 
-# embedplyr: Tools for Working With Text Embeddings
+# embedplyr <img src="man/figures/logo.png" align="right" height="138" alt="" />
 
@@
 
-## About
+## Overview
 
 embedplyr enables common operations with word and text embeddings within a 'tidyverse'/'quanteda' workflow, as demonstrated in [Data Science for Psychology: Natural Language](http://ds4psych.com).
-It includes simple functions for calculating common similarity metrics, as well as higher level functions for loading pretrained word embedding models (e.g.
-[GloVe](https://nlp.stanford.edu/projects/glove/)), applying them to words, aggregating to produce text embeddings, reducing dimensionality, and more.
+
+* `load_embeddings()` loads pretrained [GloVe](https://nlp.stanford.edu/projects/glove/), [word2vec](https://code.google.com/archive/p/word2vec/), [ConceptNet Numberbatch](https://github.com/commonsense/conceptnet-numberbatch), and [fastText](https://fasttext.cc) word embedding models from Internet sources or from your working directory
+* `embed_tokens()` returns the embedding for each token in a set of texts
+* `embed_docs()` generates text embeddings for a set of documents
+* `get_sims()` calculates row-wise similarity metrics between a set of embeddings and a given reference
+* `average_embedding()` calculates the (weighted) average of multiple embeddings
+* `reduce_dimensionality()` reduces the dimensionality of embeddings
+* `normalize()` and `normalize_rows()` normalize embeddings to the unit hypersphere
+* and more...
 

### 2024-12-02 README.Rmd
badges

@@
 <!-- badges: start -->
+[![R-CMD-check](https://github.com/rimonim/embedplyr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/rimonim/embedplyr/actions/workflows/R-CMD-check.yaml)
+[![codecov test coverage](https://codecov.io/github/rimonim/embedplyr/graph/badge.svg?token=CNJZ8R2LYZ)](https://codecov.io/github/rimonim/embedplyr)
 <!-- badges: end -->

### 2024-12-02 README.Rmd
remove local file reference

@@
 )
-options(embeddings.model.path = "~/Documents/data/embeddings")
 ```
@@
 
-glove_twitter_25d <- load_embeddings("glove.twitter.27B.25d")
+glove_twitter_25d <- load_embeddings("glove.twitter.27B.25d", save = FALSE)
 ```

### 2024-12-03 README.Rmd
fix typos

--- 2024-12-02 README.Rmd
@@
 
-embedplyr enables common operations with word and text embeddings within a 'tidyverse'/'quanteda' workflow, as demonstrated in [Data Science for Psychology: Natural Language](http://ds4psych.com).
+embedplyr enables common operations with word and text embeddings within a 'tidyverse' and/or 'quanteda' workflow, as demonstrated in [Data Science for Psychology: Natural Language](http://ds4psych.com).
 
@@
 ``` r
-# devtools package required to install quanteda from Github 
 remotes::install_github("rimonim/embedplyr") 

### 2025-01-03 README.Rmd
README formatting
add new functionality to README

@@
 * `reduce_dimensionality()` reduces the dimensionality of embeddings
+* `align_embeddings()` rotates the embeddings from one model so that they can be compared with those from another
 * `normalize()` and `normalize_rows()` normalize embeddings to the unit hypersphere
@@
 ```
-`reduce_dimensionality()` can also be used to apply the same rotation other embeddings not used to find the principle components.
+`reduce_dimensionality()` can also be used to apply the same rotation to other embeddings not used to find the principle components.
 
@@
 ```
+
+#### Align and Compare Embeddings Models
+
+Aligning separately trained embeddings can be useful for tracking semantic change over time or semantic differences between training datasets.
+It can also be useful for comparing texts across languages.
+`align_embeddings(x, y)` rotates the embeddings in `x` (and changes their dimensionality if necessary) so that they can be compared with those in `y`.
+Optionally, `matching` can be used to specify one-to-one matching between  embeddings in the two models (e.g.
+a bilingual dictionary).
+
+Once aligned, groups of embeddings can be compared using `total_dist()` or `average_sim()`. These can be used to quantify the overall distance or similarity between two parallel embedding spaces.
 

### 2025-01-05 R/load_embeddings.R
add HistWords

@@ -1,177 +1,3 @@
-#' @noRd
-supported_models <- c(
-  glove.42B.300d = "https://huggingface.co/stanfordnlp/glove/resolve/main/glove.42B.300d.zip",
-  glove.840B.300d = "https://huggingface.co/stanfordnlp/glove/resolve/main/glove.840B.300d.zip",
-  glove.6B.50d = "https://github.com/piskvorky/gensim-data/releases/download/glove-wiki-gigaword-50/glove-wiki-gigaword-50.gz",
-  glove.6B.100d = "https://github.com/piskvorky/gensim-data/releases/download/glove-wiki-gigaword-100/glove-wiki-gigaword-100.gz",
-  glove.6B.200d = "https://github.com/piskvorky/gensim-data/releases/download/glove-wiki-gigaword-200/glove-wiki-gigaword-200.gz",
-  glove.6B.300d = "https://github.com/piskvorky/gensim-data/releases/download/glove-wiki-gigaword-300/glove-wiki-gigaword-300.gz",
-  glove.twitter.27B.25d = "https://github.com/piskvorky/gensim-data/releases/download/glove-twitter-25/glove-twitter-25.gz",
-  glove.twitter.27B.50d = "https://github.com/piskvorky/gensim-data/releases/download/glove-twitter-50/glove-twitter-50.gz",
-  glove.twitter.27B.100d = "https://github.com/piskvorky/gensim-data/releases/download/glove-twitter-100/glove-twitter-100.gz",
-  glove.twitter.27B.200d = "https://github.com/piskvorky/gensim-data/releases/download/glove-twitter-200/glove-twitter-200.gz",
-  GoogleNews.vectors.negative300 = "https://github.com/piskvorky/gensim-data/releases/download/word2vec-google-news-300/word2vec-google-news-300.gz",
-  numberbatch.19.08 = "https://conceptnet.s3.amazonaws.com/downloads/2019/numberbatch/numberbatch-19.08.txt.gz",
[…]

### 2025-01-05 R/read_embeddings.R

@@ -1,0 +1,368 @@
+#' Read Embeddings From a Text or Binary File
+#'
+#' Reads GloVe text format, fastText text format, word2vec binary format,
+#' and most tabular formats (csv, tsv, etc.)
+#'
+#' @param path a file path or url
+#' @param words optional list of words for which to retrieve embeddings.
+#'
+#' @details
+#' If using a custom tabular format, the file must have tokens in the first
+#' column and numbers in all other columns.
+#'
+#' @section Value:
+#' An embeddings object (a numeric matrix with tokens as rownames)
+
+#' @export
+read_embeddings <- function(path, words = NULL) {
+	path_format <- substring(path, regexpr("\\.([[:alnum:]]+)$", path) + 1L)
+	if (path_format == "rds") {
+		# rds (possibly a matrix)
+		tmpFile = tempfile(fileext = ".rds", tmpdir = tempdir())
+		on.exit(unlink(tmpFile), add = TRUE)
+		utils::download.file(path, tmpFile)
+		out <- as.embeddings(readRDS(tmpFile))
+		if (is.character(words)) out <- predict.embeddings(out, words)
+	}else if (path_format == "bin" || grepl("\\.bin\\.", path) || grepl("word2vec", path)) {
+		# word2vec binary format
+		out <- tryCatch(
+			read_word2vec(path, words = words),
+			error = function(e) stop(
+				conditionMessage(e),
+				"\nThis file was assumed based on its name to be written in word2vec bin format. ",
+				"This assumption may have been mistaken. "
+			)
+		)
+	}else{
+		if (grepl("numberbatch", path)) use_sys <- FALSE else use_sys <- TRUE
+		out <- read_table_embeddings(path, words = words, use_sys = use_sys)
+	}
+	out
+}
+
+# check if a file is gzipped
+#' @noRd
+is_gzipped <- function(file) {
+	if (is_url(file)) {
+		# For URLs, check if the URL ends with .gz or .gzip
+		return(grepl("\\.gz(ip)?$", file, ignore.case = TRUE))
+	} else {
+		# For local files, check the file signature for gzip magic numbers
+		con <- file(file, "rb")
+		magic <- readBin(con, "raw", n = 2)
+		close(con)
+		return(identical(magic, as.raw(c(0x1f, 0x8b))))
+	}
+}
+
+#' @noRd
[…]

### 2025-01-05 tests/testthat/test-load_embeddings.R

@@ -127,7 +127,8 @@ test_that("load_embeddings reads GloVe format correctly", {
 	create_glove_file(temp_file)
 
 	# Mock supported_models to point to our local file
-	local_mocked_bindings(supported_models = c(glove.test.3d = temp_file))
+	local_mocked_bindings(supported_models = c(glove.test.3d = temp_file),
+												supported_model_names = "glove.test.3d")
 
 	# Run load_embeddings
 	embeddings_loaded <- suppressMessages( load_embeddings(model_name, dir = temp_dir, save = FALSE) )
@@ -146,7 +147,8 @@ test_that("load_embeddings reads word2vec binary format correctly", {
 	create_word2vec_bin_file(temp_file)
 
 	# Mock supported_models to point to our local file
-	local_mocked_bindings(supported_models = c(word2vec.test.3d = temp_file))
+	local_mocked_bindings(supported_models = c(word2vec.test.3d = temp_file),
+												supported_model_names = "word2vec.test.3d")
 
 	# Run load_embeddings
 	embeddings_loaded <- suppressMessages(quiet( load_embeddings(model_name, dir = temp_dir, save = FALSE) ))
@@ -187,7 +189,8 @@ test_that("load_embeddings filters words correctly when 'words' parameter is use
 	create_glove_file(temp_file)
 
 	# Mock supported_models to point to our local file
-	local_mocked_bindings(supported_models = c(glove.test.3d = temp_file))
+	local_mocked_bindings(supported_models = c(glove.test.3d = temp_file),
+												supported_model_names = "glove.test.3d")
 
 	words <- c("word1", "word3")
 
@@ -208,7 +211,8 @@ test_that("load_embeddings handles unknown file formats", {
 	writeLines("some random content", temp_file)
 
 	# Mock supported_models to point to our local file
-	local_mocked_bindings(supported_models = c(unknown.format = temp_file))
+	local_mocked_bindings(supported_models = c(unknown.format = temp_file),
+												supported_model_names = "unknown.format")
 
 	expect_error(
 		suppressMessages( load_embeddings(model_name, dir = temp_dir, save = FALSE) ),
@@ -249,7 +253,8 @@ test_that("load_embeddings reads from and writes to RDS", {
 	create_word2vec_bin_file(temp_file)
 
 	# Mock supported_models to point to our local file
-	local_mocked_bindings(supported_models = c(word2vec.test.3d = temp_file))
[…]

### 2025-01-05 README.Rmd
add HistWords to README

@@
 
-* `load_embeddings()` loads pretrained [GloVe](https://nlp.stanford.edu/projects/glove/), [word2vec](https://code.google.com/archive/p/word2vec/), [ConceptNet Numberbatch](https://github.com/commonsense/conceptnet-numberbatch), and [fastText](https://fasttext.cc) word embedding models from Internet sources or from your working directory
+* `load_embeddings()` loads pretrained [GloVe](https://nlp.stanford.edu/projects/glove/), [word2vec](https://code.google.com/archive/p/word2vec/), [HistWords](https://nlp.stanford.edu/projects/histwords/), [ConceptNet Numberbatch](https://github.com/commonsense/conceptnet-numberbatch), and [fastText](https://fasttext.cc) word embedding models from Internet sources or from your working directory
 * `embed_tokens()` returns the embedding for each token in a set of texts

### 2025-01-05 README.Rmd
Update README

@@
 
+```{r, message=FALSE, warning=FALSE}
+library(embedplyr)
+```
+
 ### Loading Pretrained Embeddings
@@
 This is especially useful for pretrained word embedding models like GloVe, word2vec, and fastText.
-Dozens of these models can be conveniently downloaded from online sources with `load_embeddings()`.
+Hundreds of these models can be conveniently downloaded from online sources with `load_embeddings()`.
 
-```{r, warning=FALSE, message=FALSE}
-library(embedplyr)
+One particularly useful feature of `load_embeddings()` is the optional `words` parameter, which allows the user to specify a subset of words to load from the model.
+This allows users to work with large models, which are often too large to load into an interactive environment in their entirety.
+For example, a user may specify the set of unique tokens in their corpus of interest, and load only these from the model.
 
+```{r, eval=FALSE, warning=FALSE, message=FALSE}
+# load 25d GloVe model trained on Twitter
 glove_twitter_25d <- load_embeddings("glove.twitter.27B.25d")
+
+# load words from 300d model trained on Google Books English Fiction 1800-1810
+eng.fiction.all_sgns.1800 <- load_embeddings(
+	"eng.fiction.all_sgns.1800",
+	words = c("word", "token", "lemma")
+	)
 ```
@@
 
-Whereas indexing a regular matrix by rownames gets slower as the number of rows increases, embedingplyr's hash table indexing means that token embeddings can be retrieved in milliseconds even from models with millions of rows.
+Whereas indexing a regular matrix by rownames gets slower as the number of rows increases, embedplyr's hash table indexing means that token embeddings can be retrieved in milliseconds even from models with millions of rows.
 
@@
 `embed_docs()` can also be used to generate other types of embeddings.
-For example, we can use the '[text](https://r-text.org)' package to generate embeddings using any model available from Huggingface transformers.
+For example, we can use the '[text](https://r-text.org)' package to generate embeddings using any model available from Hugging Face Transformers.
 

### 2025-01-08 R/sim_matrix.R
recursive behavior for sim_matrix()

@@ -1,6 +1,6 @@
 #' Pairwise Similarity or Distance Matrix
 #'
-#' Calculate a matrix of similarity or scores between the rows of the input.
+#' Calculate a matrix of similarity scores between the rows of the input.
 #'
 #' @param x an embeddings object, matrix, or dataframe with one embedding per row
 #' @param cols tidyselect - columns that contain numeric embedding values
@@ -60,6 +60,7 @@ sim_matrix_list <- list(
 #' @importFrom rlang %||%
 #' @export
 sim_matrix.default <- function(x, method = c("cosine", "cosine_squished", "euclidean", "minkowski", "dot_prod", "anchored"), ..., tidy_output = FALSE) {
+	if (inherits(x, "list")) return(lapply(x, sim_matrix))
 	if (!inherits(x, "matrix")) stop("x must be an embeddings object or numeric matrix")
 	if (is.character(method)) {
 		method_name <- method[1]

### 2025-01-08 paper/paper.Rmd
paper

---
title: 'embedplyr: Tools for Working With Text Embeddings'
tags:
  - R
  - natural language processing
  - embeddings
  - word embeddings
  - sentiment analysis
  - psychology
authors:
  - name: Louis Teitelbaum
    orcid: 0009-0001-9347-0145
    affiliation: 1
affiliations:
  - index: 1
    name: Ben-Gurion University of the Negev, Israel
date: 10 January 2025
year: 2025
bibliography: paper.bib
output: rticles::joss_article
csl: apa.csl
journal: JOSS
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(embedplyr)
```

# Aims of the Package

Dense vector embeddings are the fundamental building block of modern natural language processing [@lauriola2022]. The ability to represent the meaning of texts as a continuous, high dimensional space underlies the recent successes of large language models (LLMs). Embeddings have also revolutionized research methods and inspired new theoretical frameworks in linguistics, psychology, sociology, and neuroscience [see e.g. @duran2019; @feuerriegel2025; @grand2022; @hamilton2018; @kjell2023; @kozlowski2019; @schrimpf2021]. As text embeddings become ubiquitous in the social and behavioral sciences, the need for flexible, easy-to-learn tools increases. Answering this need, **embedplyr** (pronounced "embe-DEE-plier") enables common operations with word and text embeddings within a **tidyverse** [@wickham2019] and/or **quanteda** [@benoit2018] workflow, as demonstrated in @teitelbaum2024. **embedplyr** is designed to facilitate the use of word and text embeddings in common data manipulation and text analysis workflows, without introducing new syntax or unfamiliar data structures to R users.

Existing tools for working with embeddings in R are generally specific to particular model architectures. For example, **word2vec** [@wijffels2023] provides access to word2vec models [@mikolov2013a; @mikolov2013b], **text2vec** [@selivanov2023] provides access to LDA, LSA, and GloVe models [@blei2003; @deerwester1990; @pennington2014], **fastText** [@mouselimis2024] provides access to fastText models [@facebook2016; @bojanowski2017a; @bojanowski2017b; @grave2018], and **text** [@kjell2023] provides access to transformers-based LLMs [@wolf2020]. While these tools are already invaluable in the social and behavioral sciences, using different syntax and separate data structures for each model architecture can be cumbersome. This is especially true given the prevalence of studies that make use of multiple model architectures in parallel analyses [e.g. @carrella2023; @hussain2024; @markus2024]. **embedplyr** fills this gap with a model agnostic approach; it can be used to work with embeddings from any model framework using syntax that will be familiar to any **tidyverse** user. While some existing tools focus on convenience functions for encouraging expert-recommended best practices [e.g. @kjell2023], **embedplyr** prioritizes flexibility---much like its namesake, **dplyr** [@wickham2023]. This approach yields simple, modular functions that are useful both for educating students [see @teitelbaum2024] and experimenting with novel methods.

# Features

## Loading Pretrained Token Embeddings

**embedplyr** does not include tools for training new embedding models, but it can load embeddings from a file or download them from online. This is especially useful for pretrained word embedding models like GloVe [@pennington2014], word2vec [@mikolov2013a; @mikolov2013b], HistWords [@hamilton2018], and fastText [@bojanowski2017a]. Hundreds of these models can be conveniently downloaded from online sources with `load_embeddings()`, forgoing the need to search for model files online and juggle incompatible file types.

One particularly useful feature of `load_embeddings()` is the optional `words` parameter, which allows the user to specify a subset of words to load from the model. This allows users to work with large models, which are often too large to load into an interactive environment in their entirety. For example, a user may specify the set of unique tokens in their corpus of interest, and load only these from the model.

```{r, warning=FALSE, message=FALSE}
# load 25d GloVe model trained on Twitter
glove_twitter_25d <- load_embeddings("glove.twitter.27B.25d")
```

```{r, eval=FALSE}
# load words from model trained on Google Books English Fiction 1800-1810
eng.fiction.all_sgns.1800 <- load_embeddings(
	"eng.fiction.all_sgns.1800",
	words = c("word", "token", "lemma")
	)
```

The output of `load_embeddings()` is an embeddings object. An embeddings object is simply a numeric matrix with fast hash table indexing by rownames (generally tokens). This means that it can be easily coerced to a dataframe or tibble, while also allowing special embeddings-specific methods and functions, such as `predict.embeddings()` and `find_nearest()`:

```{r}
moral_embeddings <- predict(glove_twitter_25d, c("good", "bad"))
moral_embeddings

find_nearest(glove_twitter_25d, "dog", 5L, method = "cosine")
```

Whereas indexing a regular matrix by rownames gets slower as the number of rows increases, **embedplyr**'s hash table indexing means that token embeddings can be retrieved in milliseconds even from models with millions of rows.

## Embed Texts of Interest

Given a tidy dataframe of texts, `embed_docs()` will generate embeddings by averaging the embeddings of tokens in each text [see @ethayarajh2019; @teitelbaum2024, chap. 18]. By default, `embed_docs()` uses a simple unweighted mean, but many other averaging methods are available.

The following example embeds three texts, which for the sake of the example I will consider to have been written by a participant diagnosed with depression, one diagnosed with anxiety, and one control.

```{r, warning=FALSE, message=FALSE}
library(dplyr)

psych_df <- tribble(
    ~id,           ~text,
    "control",     "yesterday I took my dog for a walk",
    "depression",  "I slept all day and cried in the evening",
    "anxiety",     "I kept thinking of all the things I needed to do"
    )

# add embeddings to data frame
psych_embeddings_df <- psych_df |> 
    embed_docs("text", glove_twitter_25d, id_col = "id", .keep_all = TRUE)
```

`embed_tokens()` is similar to `embed_docs()`, but returns the embedding of each individual token, rather than averaging within documents.

## Embed Dictionaries

I use Distributed Dictionary Representation (DDR) as a showcase for **embedplyr**'s functionality. DDR enables the application of validated psychometric lexicons [e.g. @boyd2022] to rich, embedding-based semantic representation [@garten2018]. This is achieved by retrieving pretrained word embeddings for each word in the dictionary, and averaging them to create a single vector---the DDR. The dictionary construct can then be measured by comparing text embeddings to the DDR.

DDR is ideal for studies of abstract constructs like emotions, that refer to the general gist of a text rather than particular words. The rich representation of word embeddings allows DDR to capture even the subtlest associations between words and constructs, and to precisely reflect the *extent* to which each word is associated with each construct. It can do this even for texts that do not contain any dictionary words. Because embeddings are continuous and calibrated by design to the probabilities of word use in language, DDR also avoids difficult statistical problems that arise due to the strange distributions of word counts [see @teitelbaum2024, chap. 16].

In the following example, I construct DDRs for high and low anxiety. For the sake of the example, I use made-up dictionaries. Once embeddings are produced for each word in the dictionary, they are averaged using `average_embedding()`. By default this is a simple mean, but `average_embedding()` also supports the geometric median [@cardot2022], weighted geometric median, and weighted mean, including weighting by word frequency or smooth inverse frequency [@arora2017].

```{r, warning=FALSE}
# positive and negative construct dictionaries
high_anx_dict <- c("anxious", "overwhelmed", "nervous", "stressed")
low_anx_dict <- c("relaxed", "calm", "mellow")

# embed dictionaries
high_anx_dict_embeddings <- predict(glove_twitter_25d, high_anx_dict)
low_anx_dict_embeddings <- predict(glove_twitter_25d, low_anx_dict)

# average embeddings to create DDR
high_anx_DDR <- average_embedding(high_anx_dict_embeddings)
low_anx_DDR <- average_embedding(low_anx_dict_embeddings)
```

`average_embedding()` could be used in a similar manner to construct contextualized construct representations [@atari2023].

## Calculate Similarity Metrics

To complete the DDR analysis, I compare the embeddings of each the corpus texts to that of the DDR. This could be done by computing cosine similarity between each text and `high_anx_DDR` (`"cosine"` is the default method for `get_sims()`). In this case however, I use an anchored vector to quantify the extent to which these texts reflect high anxiety _as opposed to low anxiety_. `method = "anchored"` gives the position of each embedding on the spectrum between two anchor points, where vectors aligned with `pos` are given a score of 1 and those aligned with `neg` are given a score of 0. This approach is also known as semantic projection [@grand2022].

```{r}
anxiety_scores_df <- psych_embeddings_df |> 
  get_sims(
    dim_1:dim_25, 
    list(anxiety = list(pos = high_anx_DDR, neg = low_anx_DDR)),
    method = "anchored"
    )
anxiety_scores_df
```

Note that `get_sims()` requires only a dataframe, tibble, or embeddings object with numeric columns; the embeddings can come from any source.

## Other Functions

### Loading Other Embeddings

Embeddings can come from many sources. Accordingly, **embedplyr** provides versatile tools to read embeddings from a variety of different formats. For example, @meng2019 provide spherical text embeddings pretrained on Wikipedia, formatted as text files. After downloading one of these files, it can be loaded using `read_embeddings()`. `read_embeddings()` can read GloVe text format, fastText text format, word2vec binary format, and most tabular formats (csv, tsv, etc.).

```{r, eval=FALSE}
spherical_mod <- read_embeddings("~/Downloads/jose_50d.txt")
```

Various R data types can also be converted to embeddings objects using `as.embeddings()`.

### Arbitrary Embedding Functions

In addition to embedding documents by averaging static word embeddings, `embed_docs()` can be used to wrap arbitrary embedding functions. For example, the **text** package [@kjell2023] can be used to generate embeddings using any model available from Hugging Face Transformers [@wolf2020]. This ability to wrap arbitrary functions is useful for analyses which require multiple corpora to be embedded in the same manner.

```{r, eval=FALSE}
# add embeddings to data frame
valence_sbert_df <- valence_df |> 
	embed_docs("text", text::textEmbed, id_col = "id", .keep_all = TRUE)
```

### Quanteda Compatibility

The examples above used a **tidyverse** approach with tibbles as the primary data structure, but **embedplyr** functions can be used similarly within a **quanteda** workflow. In the example below, I reproduce the text embedding shown above using `textstat_embedding()` (following **quanteda** naming standards) as opposed to `embed_docs()`.

```{r, warning=FALSE, message=FALSE}
library(quanteda)

# corpus
psych_embeddings_corp <- corpus(psych_embeddings_df, docid_field = "id")

# dfm
psych_embeddings_dfm <- psych_embeddings_corp |> 
	tokens() |> 
	dfm()

# compute embeddings
psych_embeddings_df <- psych_embeddings_dfm |> 
	textstat_embedding(glove_twitter_25d)
```

### Similarity Metrics

Functions for similarity and distance metrics are as simple as possible; each one takes in vectors and outputs a scalar. In addition to cosine similarity, Euclidean distance, and dot product, **embedplyr** functions also support projection to anchored vectors [@grand2022], as shown in the examples above.

```{r, eval=FALSE}
dot_product <- dot_prod(vec1, vec2)
cosine_similarity <- cos_sim(vec1, vec2)
euclidean_distance <- euc_dist(vec1, vec2)
semantic_projection <- anchored_sim(vec1, pos = vec2, neg = vec3)
```

### Reduce Dimensionality

It is sometimes useful to reduce the dimensionality of embeddings. This is done with `reduce_dimensionality()`, which by default performs PCA without column normalization. Reducing the dimensionality to two or three dimensions can be useful for visualization.

```{r}
valence_df_2d <- psych_embeddings_df |> 
	reduce_dimensionality(dim_1:dim_25, 2)
valence_df_2d
```

`reduce_dimensionality()` can also be used to apply the same rotation to other embeddings not used to find the principle components.

```{r}
new_embeddings <- predict(glove_twitter_25d, c("new", "strange"))

# get rotation with `output_rotation = TRUE`
valence_rotation_2d <- psych_embeddings_df |> 
	reduce_dimensionality(dim_1:dim_25, 2, output_rotation = TRUE)

# apply the same rotation to new embeddings
new_with_valence_rotation <- new_embeddings |> 
	reduce_dimensionality(custom_rotation = valence_rotation_2d)
new_with_valence_rotation
```

### Align and Compare Embeddings Models

Aligning separately trained embeddings can be useful for tracking semantic change over time or semantic differences between training datasets. It can also be useful for comparing texts across languages. `align_embeddings(x, y)` rotates the embeddings in `x` (and changes their dimensionality if necessary) so that they can be compared with those in `y` [@schonemann1966]. Optionally, `matching` can be used to specify one-to-one matching between  embeddings in the two models (e.g. a bilingual dictionary).

Once aligned, groups of embeddings can be compared using `total_dist()` or `average_sim()`. These can be used to quantify the overall distance or similarity between two parallel embedding spaces.

### Normalize (Scale Embeddings to the Unit Hypersphere)

`normalize()` and `normalize_rows()` scale embeddings such that their magnitude is 1, while their angle from the origin is unchanged.

```{r}
normalize(1:10)

normalize(moral_embeddings)

psych_embeddings_normalized <- psych_embeddings_df |>
	normalize_rows(dim_1:dim_25)
```

### Magnitude

The magnitude, norm, or length of a vector is its Euclidean distance from the origin.

```{r}
magnitude(1:10)

magnitude(moral_embeddings)
```

### Similarity Matrix

Given an embeddings object or data frame of embeddings, `sim_matrix()` calculates a matrix of similarity scores between the rows of the input. This can be useful for analyzing distributional properties of groups of documents [see e.g. @markus2024] or of groups of tokens [see e.g. @choi2024].

```{r, warning=FALSE}
text <- c("lions and tigers and bears", "curious green ideas")

text |> 
	embed_tokens(glove_twitter_25d, output_embeddings = TRUE) |> 
	sim_matrix(method = "euclidean")
```

# Licensing and Availability

**embedplyr** is licensed under the GNU General Public License (v3.0). All of its source code is stored publicly on Github (https://github.com/rimonim/embedplyr), with a corresponding issue tracker.

### 2025-01-08 paper/paper.Rmd
compile paper workflow

@@
 bibliography: paper.bib
-output: rticles::joss_article
+output:
+  html_document:
+    df_print: paged
 csl: apa.csl

### 2025-01-14 README.Rmd
minor revisions

@@
 ```{r, eval=FALSE}
-# function that takes character vector and outputs a data frame
-sbert_embeddings <- function(texts) {
-	text::textEmbed(
-		texts,
-		model = "sentence-transformers/all-MiniLM-L12-v2", # model name
-		layers = -2,  # second to last layer (default)
-		tokens_select = "[CLS]", # use only [CLS] token
-		dim_name = FALSE,
-		keep_token_embeddings = FALSE
-	)$texts[[1]]
-}
-
+# add embeddings to data frame
 valence_sbert_df <- valence_df |> 
-	embed_docs("text", sbert_embeddings, id_col = "id", .keep_all = TRUE)
+	embed_docs(
+		"text", text::textEmbed, id_col = "id", .keep_all = TRUE,
+		model = "sentence-transformers/all-MiniLM-L12-v2"
+		)
 ```

### 2025-01-14 paper/paper.Rmd
minor revisions

@@
 
-The following example embeds three texts, which for the sake of the example I will consider to have been written by a participant diagnosed with depression, one diagnosed with anxiety, and one control.
+The following example embeds three texts, which for the sake of the example will can considered to have been written by one participant diagnosed with depression, one diagnosed with anxiety, and one control.
 
@@
 
-I use Distributed Dictionary Representation (DDR) as a showcase for **embedplyr**'s functionality.
-DDR enables the application of validated psychometric lexicons [e.g.
+Distributed Dictionary Representation (DDR) enables the application of validated psychometric lexicons [e.g.
 @boyd2022] to rich, embedding-based semantic representation [@garten2018].
@@
 
-In the following example, I construct DDRs for high and low anxiety.
-For the sake of the example, I use made-up dictionaries.
+In the following example, DDRs are constructed for high and low anxiety using example dictionaries.
 Once embeddings are produced for each word in the dictionary, they are averaged using `average_embedding()`.
@@
 
-To complete the DDR analysis, I compare the embeddings of each the corpus texts to that of the DDR.
+To complete the DDR analysis initiated above, the embeddings of each the corpus texts are compared to that of the DDR.
 This could be done by computing cosine similarity between each text and `high_anx_DDR` (`"cosine"` is the default method for `get_sims()`).
-In this case however, I use an anchored vector to quantify the extent to which these texts reflect high anxiety _as opposed to low anxiety_.
+In this case however, an anchored vector is used to quantify the extent to which these texts reflect high anxiety _as opposed to low anxiety_.
 `method = "anchored"` gives the position of each embedding on the spectrum between two anchor points, where vectors aligned with `pos` are given a score of 1 and those aligned with `neg` are given a score of 0.
@@
 
-### Loading Other Embeddings
+### Read Embeddings From File
 
@@
 valence_sbert_df <- valence_df |> 
-	embed_docs("text", text::textEmbed, id_col = "id", .keep_all = TRUE)
+	embed_docs(
+		"text", text::textEmbed, id_col = "id", .keep_all = TRUE,
+		model = "sentence-transformers/all-MiniLM-L12-v2"
+		)
 ```

### 2025-01-14 R/load_embeddings.R

@@ -1,8 +1,8 @@
 #' Load Pretrained GloVe, word2vec, and fastText Embeddings
 #'
 #' Loads pretrained word embeddings. If the specified model has already been
-#' downloaded, it is read from file with [read_embeddings()]. If not, the model
-#' is retrieved from online sources and, by default, saved.
+#' downloaded, it is read from file with a call to [read_embeddings()]. If not,
+#' the model is retrieved from online sources and, by default, saved.
 #'
 #' @param model the name of a supported model
 #' @param dir directory in which the model is or should be saved when `save = TRUE`.

### 2025-02-11 paper/paper.Rmd
shorten paper

@@
 
-# Aims of the Package
+# Statement of Need
 
@@
 This allows users to work with large models, which are often too large to load into an interactive environment in their entirety.
-For example, a user may specify the set of unique tokens in their corpus of interest, and load only these from the model.
 
@@
 
-DDR is ideal for studies of abstract constructs like emotions, that refer to the general gist of a text rather than particular words.
-The rich representation of word embeddings allows DDR to capture even the subtlest associations between words and constructs, and to precisely reflect the *extent* to which each word is associated with each construct.
-It can do this even for texts that do not contain any dictionary words.
-Because embeddings are continuous and calibrated by design to the probabilities of word use in language, DDR also avoids difficult statistical problems that arise due to the strange distributions of word counts [see @teitelbaum2024, chap.
-16].
-
 In the following example, DDRs are constructed for high and low anxiety using example dictionaries.
@@
 
-## Other Functions
-
-### Read Embeddings From File
-
-Embeddings can come from many sources.
-Accordingly, **embedplyr** provides versatile tools to read embeddings from a variety of different formats.
-For example, @meng2019 provide spherical text embeddings pretrained on Wikipedia, formatted as text files.
-After downloading one of these files, it can be loaded using `read_embeddings()`.
-`read_embeddings()` can read GloVe text format, fastText text format, word2vec binary format, and most tabular formats (csv, tsv, etc.).
-
-```{r, eval=FALSE}
-spherical_mod <- read_embeddings("~/Downloads/jose_50d.txt")
-```
-
-Various R data types can also be converted to embeddings objects using `as.embeddings()`.
-
-### Arbitrary Embedding Functions
-
-In addition to embedding documents by averaging static word embeddings, `embed_docs()` can be used to wrap arbitrary embedding functions.
-For example, the **text** package [@kjell2023] can be used to generate embeddings using any model available from Hugging Face Transformers [@wolf2020].
-This ability to wrap arbitrary functions is useful for analyses which require multiple corpora to be embedded in the same manner.
-
-```{r, eval=FALSE}
-# add embeddings to data frame
-valence_sbert_df <- valence_df |> 
-	embed_docs(
-		"text", text::textEmbed, id_col = "id", .keep_all = TRUE,
-		model = "sentence-transformers/all-MiniLM-L12-v2"
-		)
-```
-
-### Quanteda Compatibility
-
-The examples above used a **tidyverse** approach with tibbles as the primary data structure, but **embedplyr** functions can be used similarly within a **quanteda** workflow.
-In the example below, I reproduce the text embedding shown above using `textstat_embedding()` (following **quanteda** naming standards) as opposed to `embed_docs()`.
-
-```{r, warning=FALSE, message=FALSE}
-library(quanteda)
-
-# corpus
-psych_embeddings_corp <- corpus(psych_embeddings_df, docid_field = "id")
-
-# dfm
-psych_embeddings_dfm <- psych_embeddings_corp |> 
-	tokens() |> 
-	dfm()
-
-# compute embeddings
-psych_embeddings_df <- psych_embeddings_dfm |> 
-	textstat_embedding(glove_twitter_25d)
-```
-
-### Similarity Metrics
-
-Functions for similarity and distance metrics are as simple as possible; each one takes in vectors and outputs a scalar.
-In addition to cosine similarity, Euclidean distance, and dot product, **embedplyr** functions also support projection to anchored vectors [@grand2022], as shown in the examples above.
-
-```{r, eval=FALSE}
-dot_product <- dot_prod(vec1, vec2)
-cosine_similarity <- cos_sim(vec1, vec2)
-euclidean_distance <- euc_dist(vec1, vec2)
-semantic_projection <- anchored_sim(vec1, pos = vec2, neg = vec3)
-```
-
-### Reduce Dimensionality
-
-It is sometimes useful to reduce the dimensionality of embeddings.
-This is done with `reduce_dimensionality()`, which by default performs PCA without column normalization.
-Reducing the dimensionality to two or three dimensions can be useful for visualization.
-
-```{r}
-valence_df_2d <- psych_embeddings_df |> 
-	reduce_dimensionality(dim_1:dim_25, 2)
-valence_df_2d
-```
-
-`reduce_dimensionality()` can also be used to apply the same rotation to other embeddings not used to find the principle components.
-
-```{r}
-new_embeddings <- predict(glove_twitter_25d, c("new", "strange"))
-
-# get rotation with `output_rotation = TRUE`
-valence_rotation_2d <- psych_embeddings_df |> 
-	reduce_dimensionality(dim_1:dim_25, 2, output_rotation = TRUE)
-
-# apply the same rotation to new embeddings
-new_with_valence_rotation <- new_embeddings |> 
-	reduce_dimensionality(custom_rotation = valence_rotation_2d)
-new_with_valence_rotation
-```
-
-### Align and Compare Embeddings Models
-
-Aligning separately trained embeddings can be useful for tracking semantic change over time or semantic differences between training datasets.
-It can also be useful for comparing texts across languages.
-`align_embeddings(x, y)` rotates the embeddings in `x` (and changes their dimensionality if necessary) so that they can be compared with those in `y` [@schonemann1966].
-Optionally, `matching` can be used to specify one-to-one matching between  embeddings in the two models (e.g.
-a bilingual dictionary).
-
-Once aligned, groups of embeddings can be compared using `total_dist()` or `average_sim()`. These can be used to quantify the overall distance or similarity between two parallel embedding spaces.
-
-### Normalize (Scale Embeddings to the Unit Hypersphere)
-
-`normalize()` and `normalize_rows()` scale embeddings such that their magnitude is 1, while their angle from the origin is unchanged.
-
-```{r}
-normalize(1:10)
-
-normalize(moral_embeddings)
-
-psych_embeddings_normalized <- psych_embeddings_df |>
-	normalize_rows(dim_1:dim_25)
-```
-
-### Magnitude
-
-The magnitude, norm, or length of a vector is its Euclidean distance from the origin.
-
-```{r}
-magnitude(1:10)
-
-magnitude(moral_embeddings)
-```
-
-### Similarity Matrix
-
-Given an embeddings object or data frame of embeddings, `sim_matrix()` calculates a matrix of similarity scores between the rows of the input.
-This can be useful for analyzing distributional properties of groups of documents [see e.g.
-@markus2024] or of groups of tokens [see e.g.
-@choi2024].
-
-```{r, warning=FALSE}
-text <- c("lions and tigers and bears", "curious green ideas")
-
-text |> 
-	embed_tokens(glove_twitter_25d, output_embeddings = TRUE) |> 
-	sim_matrix(method = "euclidean")
-```
-
 # Licensing and Availability

### 2025-08-05 README.Rmd
performance vignette
addresses issue #2

@@
 
-Whereas indexing a regular matrix by rownames gets slower as the number of rows increases, embedplyr's hash table indexing means that token embeddings can be retrieved in milliseconds even from models with millions of rows.
+Whereas indexing a regular matrix by rownames gets slower as the number of rows increases, embedplyr's hash table indexing means that token embeddings can be retrieved in milliseconds even from models with millions of rows (see the [performance vignette](vignettes/performance.html)).
 

### 2025-08-05 paper/paper.Rmd
performance vignette
addresses issue #2

@@
 
-Whereas indexing a regular matrix by rownames gets slower as the number of rows increases, **embedplyr**'s hash table indexing means that token embeddings can be retrieved in milliseconds even from models with millions of rows.
+Whereas indexing a regular matrix by rownames gets slower as the number of rows increases, **embedplyr**'s hash table indexing means that token embeddings can be retrieved in milliseconds even from models with millions of rows (see the [performance vignette](vignettes/performance.html)).
 

### 2025-08-06 R/embeddings-subsetting.R
consistency in linking to reference documentation
addresses issue #9

@@ -12,7 +12,7 @@
 #' @inheritParams base::subset
 #'
 #' @details
-#' The difference between `embeddings[i,]` and `predict(embeddings, i)` is that
+#' The difference between `embeddings[i,]` and [`predict(embeddings, i)`][predict.embeddings] is that
 #' the former will throw an error when items of `i` are not valid indices, whereas
 #' the latter will handle it gracefully (at the cost of a few more milliseconds
 #' if `i` is long).

### 2025-08-06 paper/paper.Rmd

@@
 Answering this need, **embedplyr** (pronounced "embe-DEE-plier") enables common operations with word and text embeddings within a **tidyverse** [@wickham2019] and/or **quanteda** [@benoit2018] workflow, as demonstrated in @teitelbaum2024.
-**embedplyr** is designed to facilitate the use of word and text embeddings in common data manipulation and text analysis workflows, without introducing new syntax or unfamiliar data structures to R users.
+Much as **dplyr** is "a grammar of data manipulation, providing a consistent set of verbs that help you solve the most common data manipulation challenges" [@wickham2025], **embedplyr** is a grammar of embeddings manipulation, designed to facilitate the use of word and text embeddings in common analysis workflows without introducing new syntax or unfamiliar data structures to R users.
 

### 2025-09-01 README.Rmd
consistency in linking to reference documentation
addresses issue #9
change `predict.embeddings()` to `emb()`
addresses issue #6

@@
 An embeddings object is just a numeric matrix with fast hash table indexing by rownames (generally tokens).
-This means that it can be easily coerced to a dataframe or tibble, while also allowing special embeddings-specific methods and functions, such as `predict.embeddings()` and `find_nearest()`:
+This means that it can be easily coerced to a dataframe or tibble, while also allowing special embeddings-specific methods and functions, such as `emb()` and `find_nearest()`:
 
 ```{r}
-moral_embeddings <- predict(glove_twitter_25d, c("good", "bad"))
+moral_embeddings <- emb(glove_twitter_25d, c("good", "bad"))
 moral_embeddings
@@
 ```{r}
-good_vec <- predict(glove_twitter_25d, "good")
-intense_vec <- predict(glove_twitter_25d, "intense")
+good_vec <- emb(glove_twitter_25d, "good")
+intense_vec <- emb(glove_twitter_25d, "intense")
 valence_quantified <- valence_embeddings_df |> 
@@
 ```{r}
-new_embeddings <- predict(glove_twitter_25d, c("new", "strange"))
+new_embeddings <- emb(glove_twitter_25d, c("new", "strange"))
 
@@
 a bilingual dictionary).
+See `align_embeddings()` for full documentation.
 

### 2025-09-01 paper/paper.Rmd
change `predict.embeddings()` to `emb()`
addresses issue #6

@@
 An embeddings object is simply a numeric matrix with fast hash table indexing by rownames (generally tokens).
-This means that it can be easily coerced to a dataframe or tibble, while also allowing special embeddings-specific methods and functions, such as `predict.embeddings()` and `find_nearest()`:
+This means that it can be easily coerced to a dataframe or tibble, while also allowing special embeddings-specific methods and functions, such as `emb()` and `find_nearest()`:
 
 ```{r}
-moral_embeddings <- predict(glove_twitter_25d, c("good", "bad"))
+moral_embeddings <- emb(glove_twitter_25d, c("good", "bad"))
 moral_embeddings
@@
 # embed dictionaries
-high_anx_dict_embeddings <- predict(glove_twitter_25d, high_anx_dict)
-low_anx_dict_embeddings <- predict(glove_twitter_25d, low_anx_dict)
+high_anx_dict_embeddings <- emb(glove_twitter_25d, high_anx_dict)
+low_anx_dict_embeddings <- emb(glove_twitter_25d, low_anx_dict)
 

### 2025-09-01 R/average_embedding.R

@@ -56,7 +56,7 @@ make_embedding_weights <- function(x_names, w){
 #' @importFrom stats weighted.mean
 #' @examples
 #' happy_dict <- c("happy", "joy", "smile", "enjoy")
-#' happy_dict_embeddings <- predict(glove_twitter_25d, happy_dict)
+#' happy_dict_embeddings <- emb(glove_twitter_25d, happy_dict)
 #'
 #' happy_dict_vec <- average_embedding(happy_dict_embeddings)
 #' happy_dict_vec_weighted <- average_embedding(happy_dict_embeddings, w = "trillion_word")

### 2025-09-01 R/emb.R

@@ -21,12 +21,11 @@
 #' @examples
 #' words <- c("happy", "sad")
 #'
-#' texts_embeddings <- predict(glove_twitter_25d, words)
+#' texts_embeddings <- emb(glove_twitter_25d, words)
 #' texts_embeddings
 
-#' @rdname predict.embeddings
 #' @export
-predict.embeddings <- function(object, newdata, drop = TRUE, .keep_missing = FALSE){
+emb <- function(object, newdata, drop = TRUE, .keep_missing = FALSE){
   if (any(zchars <- !nzchar(newdata))) {
     warning(sprintf('Replacing %d empty strings with " ".', sum(zchars)))
     newdata[zchars] <- " "

### 2025-09-01 R/embed_docs.R

@@ -115,7 +115,7 @@ textstat_embedding <- function(dfm, model, w = NULL, method = "mean", output_emb
     dfm <- suppressWarnings( quanteda::dfm_weight(dfm, weights = w) )
   }
   # find word embeddings
-  feat_embeddings <- predict.embeddings(model, feats, .keep_missing = TRUE)
+  feat_embeddings <- emb(model, feats, .keep_missing = TRUE)
 
   if (method == "median") {
     out_list <- lapply(seq_len(nrow(dfm)), function(r){

### 2025-09-01 R/embed_tokens.R

@@ -62,7 +62,7 @@ embed_tokens.default <- function(x, model, ..., .keep_missing = FALSE,
 	if (any(embedding_not_found)) {
 		warning(sprintf("%d tokens in `x` are not present in `model`.", sum(embedding_not_found)))
 	}
-	out <- lapply(x, function(nd) suppressWarnings(predict.embeddings(model, nd, drop = FALSE, .keep_missing = .keep_missing)))
+	out <- lapply(x, function(nd) suppressWarnings(emb(model, nd, drop = FALSE, .keep_missing = .keep_missing)))
 	if (output_embeddings) return(out)
 	out <- lapply(out, function(emb) if (nrow(emb) != 0) as_tibble(emb, rownames = "token") else NULL)
 	dplyr::bind_rows(out, .id = "doc_id")

### 2025-09-01 R/embeddings-subsetting.R

@@ -12,7 +12,7 @@
 #' @inheritParams base::subset
 #'
 #' @details
-#' The difference between `embeddings[i,]` and [`predict(embeddings, i)`][predict.embeddings] is that
+#' The difference between `embeddings[i,]` and [`emb(embeddings, i)`][emb] is that
 #' the former will throw an error when items of `i` are not valid indices, whereas
 #' the latter will handle it gracefully (at the cost of a few more milliseconds
 #' if `i` is long).

### 2025-09-01 R/find_nearest.R

@@ -33,7 +33,7 @@
 #'
 #' @examples
 #' words <- c("happy", "sad")
-#' words_embeddings <- predict(glove_twitter_25d, words)
+#' words_embeddings <- emb(glove_twitter_25d, words)
 #'
 #' find_nearest(glove_twitter_25d, words)
 #' find_nearest(glove_twitter_25d, words_embeddings) # equivalent to previous

### 2025-09-01 R/get_sims.R

@@ -45,9 +45,9 @@
 #'
 #' @import tibble
 #' @examples
-#' valence_embeddings <- predict(glove_twitter_25d, c("good", "bad"))
-#' happy_vec <- predict(glove_twitter_25d, "happy")
-#' sad_vec <- predict(glove_twitter_25d, "sad")
+#' valence_embeddings <- emb(glove_twitter_25d, c("good", "bad"))
+#' happy_vec <- emb(glove_twitter_25d, "happy")
+#' sad_vec <- emb(glove_twitter_25d, "sad")
 #'
 #' valence_embeddings |>
 #'   get_sims(list(happy = happy_vec))

### 2025-09-01 R/load_embeddings.R

@@ -175,7 +175,7 @@ load_embeddings <- function(model, dir = NULL, words = NULL, save = TRUE, format
     message("Reading model from file...")
     if (!new_download && saved_format == "rds") {
       out <- readRDS(model_path)
-      if (is.character(words)) out <- predict.embeddings(out, words)
+      if (is.character(words)) out <- emb(out, words)
     }else{
       out <- read_embeddings(model_path, words)
     }

### 2025-09-01 R/magnitude.R

@@ -13,7 +13,7 @@
 #' vec <- c(1, 4, 2)
 #' magnitude(vec)
 #'
-#' valence_embeddings <- predict(glove_twitter_25d, c("good", "bad"))
+#' valence_embeddings <- emb(glove_twitter_25d, c("good", "bad"))
 #' magnitude(valence_embeddings)
 #'
 #' embeddings_list <- find_nearest(glove_twitter_25d, c("good", "bad"), each = TRUE)

### 2025-09-01 R/normalize.R

@@ -15,7 +15,7 @@
 #' vec <- c(1, 4, 2)
 #' normalize(vec)
 #'
-#' valence_embeddings <- predict(glove_twitter_25d, c("good", "bad"))
+#' valence_embeddings <- emb(glove_twitter_25d, c("good", "bad"))
 #' normalize(valence_embeddings)
 #'
 #' embeddings_list <- find_nearest(glove_twitter_25d, c("good", "bad"), each = TRUE)

### 2025-09-01 R/read_embeddings.R

@@ -22,7 +22,7 @@ read_embeddings <- function(path, words = NULL) {
 		on.exit(unlink(tmpFile), add = TRUE)
 		utils::download.file(path, tmpFile)
 		out <- as.embeddings(readRDS(tmpFile))
-		if (is.character(words)) out <- predict.embeddings(out, words)
+		if (is.character(words)) out <- emb(out, words)
 	}else if (path_format == "bin" || grepl("\\.bin\\.", path) || grepl("word2vec", path)) {
 		# word2vec binary format
 		out <- tryCatch(

### 2025-09-01 R/sim_matrix.R

@@ -34,7 +34,7 @@
 #'
 #' @import tibble
 #' @examples
-#' emb <- predict(glove_twitter_25d, c("table", "chair", "cat"))
+#' emb <- emb(glove_twitter_25d, c("table", "chair", "cat"))
 #'
 #' sim_matrix(emb)
 #' sim_matrix(emb, method = "euclidean")

### 2025-09-01 R/vector_utilities.R

@@ -11,12 +11,12 @@
 #'
 #' @examples
 #' emotion_words <- c("happy", "sad", "scared", "angry", "anxious", "grateful")
-#' emotion_embeddings <- predict(glove_twitter_25d, emotion_words)
+#' emotion_embeddings <- emb(glove_twitter_25d, emotion_words)
 #' emotion_embeddings_2d <- reduce_dimensionality(emotion_embeddings, 2)
 #'
 #' # project points
-#' happy_vec_2d <- predict(emotion_embeddings_2d, "happy")
-#' sad_vec_2d <- predict(emotion_embeddings_2d, "sad")
+#' happy_vec_2d <- emb(emotion_embeddings_2d, "happy")
+#' sad_vec_2d <- emb(emotion_embeddings_2d, "sad")
 #' emotional_projections <- emotion_embeddings_2d |>
 #'   project_points_onto_line(happy_vec_2d, sad_vec_2d)
 #'

### 2025-09-01 data-raw/glove_twitter_25d.R

@@ -3,6 +3,6 @@ devtools::load_all()
 options(embeddings.model.path = "~/Documents/data/embeddings")
 glove_twitter_25d <- load_embeddings("glove.twitter.27B.25d")
 top_words <- trillion_word[1:12000]
-glove_twitter_25d <- predict(glove_twitter_25d, names(top_words))
+glove_twitter_25d <- emb(glove_twitter_25d, names(top_words))
 
 usethis::use_data(glove_twitter_25d, overwrite = TRUE)

### 2025-09-01 tests/testthat/test-emb.R

@@ -4,11 +4,11 @@ create_sample_embeddings <- function() {
 						 dimnames = list(tokens, paste0("dim_", 1:5)))
 }
 
-test_that("predict.embeddings handles missing tokens with .keep_missing = FALSE", {
+test_that("emb handles missing tokens with .keep_missing = FALSE", {
 	embeddings_obj <- create_sample_embeddings()
 	newdata <- c("happy", "unknown", "sad")
 	expect_warning(
-		result <- predict(embeddings_obj, newdata, .keep_missing = FALSE),
+		result <- emb(embeddings_obj, newdata, .keep_missing = FALSE),
 		"1 items in `newdata` are not present in the embeddings object."
 	)
 	expect_true(is.embeddings(result))
@@ -16,11 +16,11 @@ test_that("predict.embeddings handles missing tokens with .keep_missing = FALSE"
 	expect_equal(rownames(result), c("happy", "sad"))
 })
 
-test_that("predict.embeddings handles missing tokens with .keep_missing = TRUE", {
+test_that("emb handles missing tokens with .keep_missing = TRUE", {
 	embeddings_obj <- create_sample_embeddings()
 	newdata <- c("happy", "unknown", "sad")
 	expect_warning(
-		result <- predict(embeddings_obj, newdata, .keep_missing = TRUE),
+		result <- emb(embeddings_obj, newdata, .keep_missing = TRUE),
 		"1 items in `newdata` are not present in the embeddings object."
 	)
 	expect_true(is.embeddings(result))
@@ -31,21 +31,21 @@ test_that("predict.embeddings handles missing tokens with .keep_missing = TRUE",
 	expect_true(all(is.na(result["unknown", ])))
 })
 
-test_that("predict.embeddings handles duplicate tokens in newdata", {
+test_that("emb handles duplicate tokens in newdata", {
 	embeddings_obj <- create_sample_embeddings()
 	newdata <- c("happy", "sad", "happy")
-	result <- predict(embeddings_obj, newdata)
+	result <- emb(embeddings_obj, newdata)
 	expect_true(is.embeddings(result))
 	expect_equal(nrow(result), 3)
 	expect_equal(rownames(result), newdata)
 	expect_equal(result[1, ], result[3, ])
 })
 
-test_that("predict.embeddings handles empty strings in newdata", {
+test_that("emb handles empty strings in newdata", {
 	embeddings_obj <- create_sample_embeddings()
 	newdata <- c("happy", "", "sad")
[…]

### 2025-09-01 tests/testthat/test-load_embeddings.R

@@ -105,8 +105,8 @@ test_that("load_embeddings loads from Internet", {
 	skip_if_not(curl::has_internet())
 	# without save
 	glove_test <- quiet(suppressMessages(load_embeddings("glove.6B.50d", save = FALSE)))
-	red_blue <- cos_sim(predict(glove_test, "red"), predict(glove_test, "blue"))
-	red_dimension <- cos_sim(predict(glove_test, "red"), predict(glove_test, "dimension"))
+	red_blue <- cos_sim(emb(glove_test, "red"), emb(glove_test, "blue"))
+	red_dimension <- cos_sim(emb(glove_test, "red"), emb(glove_test, "dimension"))
 	expect_true(red_blue > red_dimension)
 	# with save and words
 	expect_warning(
@@ -114,7 +114,7 @@ test_that("load_embeddings loads from Internet", {
 		"Saving only a subset of words is not supported for original format"
 		)
 	expect_equal(nrow(glove_test), 3)
-	expect_true(red_blue == cos_sim(predict(glove_test, "red"), predict(glove_test, "blue")))
+	expect_true(red_blue == cos_sim(emb(glove_test, "red"), emb(glove_test, "blue")))
 })
 
 # CRAN-friendly tests

### 2025-09-29 README.Rmd
Refrence contributor guidelines in README
addresses JOSS reviewer feedback #1

@@
 ```
+
+## Contributor Guidelines
+
+For issues, bug reports, and feature requests, [file a GitHub issue](https://github.com/rimonim/embedplyr/issues).
+For support or other inquiries, feel free to contact the maintainer by email at [info@ds4psych.com](mailto:info@ds4psych.com).
+However, the best way to get something changed is to do it yourself.
+We welcome contributions!
+See the [contributor guidelines](https://github.com/rimonim/embedplyr/wiki/Contributor-Guidelines) for details.

### 2025-10-10 paper/paper.Rmd
add summary

@@
 bibliography: paper.bib
-output:
-  html_document:
-    df_print: paged
+output: pdf
 csl: apa.csl
@@
 
-# Statement of Need
+# Summary
 
@@
 As text embeddings become ubiquitous in the social and behavioral sciences, the need for flexible, easy-to-learn tools increases.
-Answering this need, **embedplyr** (pronounced "embe-DEE-plier") enables common operations with word and text embeddings within a **tidyverse** [@wickham2019] and/or **quanteda** [@benoit2018] workflow, as demonstrated in @teitelbaum2024.
+Answering this need, **embedplyr** (pronounced "embe-DEE-plier") enables common operations with word and text embeddings within familiar analysis workflows.
+
+# Statement of Need
+
+**embedplyr** is designed for integration with a **tidyverse** [@wickham2019] and/or **quanteda** [@benoit2018] workflow, as demonstrated in @teitelbaum2024.
 Much as **dplyr** is "a grammar of data manipulation, providing a consistent set of verbs that help you solve the most common data manipulation challenges" [@wickham2025], **embedplyr** is a grammar of embeddings manipulation, designed to facilitate the use of word and text embeddings in common analysis workflows without introducing new syntax or unfamiliar data structures to R users.
@@
 @carrella2023; @hussain2024; @markus2024].
-**embedplyr** fills this gap with a model agnostic approach; it can be used to work with embeddings from any model framework using syntax that will be familiar to any **tidyverse** user.
+**embedplyr** fills this gap with a model-agnostic approach; it can be used to work with embeddings from any model framework using syntax that will be familiar to any **tidyverse** user.
 While some existing tools focus on convenience functions for encouraging expert-recommended best practices [e.g.
@@
 
-The following example embeds three texts, which for the sake of the example will can considered to have been written by one participant diagnosed with depression, one diagnosed with anxiety, and one control.
+The following example embeds three texts, which for the sake of the example can be considered to have been written by one participant diagnosed with depression, one diagnosed with anxiety, and one control.
 
@@
 **embedplyr** is licensed under the GNU General Public License (v3.0). All of its source code is stored publicly on Github (https://github.com/rimonim/embedplyr), with a corresponding issue tracker.
+
+# References

### 2025-11-10 paper/paper.Rmd
references formatting
update acknowledgements

@@
 While some existing tools focus on convenience functions for encouraging expert-recommended best practices [e.g.
-@kjell2023], **embedplyr** prioritizes flexibility---much like its namesake, **dplyr** [@wickham2023].
+@kjell2023], **embedplyr** prioritizes flexibility---much like its namesake, **dplyr** [@wickham2025].
 This approach yields simple, modular functions that are useful both for educating students [see @teitelbaum2024] and experimenting with novel methods.
@@
 
+# Acknowledgements
+
+The author thanks Almog Simchon for helpful feedback in the early stages of development. This work was supported in part by a grant by the Ministry of Aliyah and Integration of the State of Israel.
+
 # References

### 2025-11-11 README.Rmd
JOSS publication

@@
 [![codecov test coverage](https://codecov.io/github/rimonim/embedplyr/graph/badge.svg?token=CNJZ8R2LYZ)](https://codecov.io/github/rimonim/embedplyr)
+[![DOI](https://joss.theoj.org/papers/10.21105/joss.08262/status.svg)](https://doi.org/10.21105/joss.08262)
 <!-- badges: end -->
